---
title: "Airline Customer Satisfaction - Group 17"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
    css: styles.css
  pdf_document: default
date: "2025-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analyst {.unnumbered .no-toc}

| Name                   | Student Number |
|------------------------|----------------|
| Andra Miftah Ar Rahman | 5728827        |
| Vicky Wijaya           | 5795427        |
| Maulvi Azmi Winata     | 6146015        |
| Dewi Farra Prasasya    | 6171486        |
| Agassi Fathurisma      | 6211542        |

# Project Background <a name="1"></a>

## Abstract <a name="1.1"></a>

**Framework for abstract:**

After industry-wide setbacks from the COVID-19 pandemic, the airline industry is experiencing a strong recovery in profitability. According to the International Air Transport Association (IATA), airlines worldwide have seen improved projections since 2024. The market is projected to showcase an annual growth rate (CAGR 2025-2029) of 4.36%, resulting in a projected market volume of US\$771.26bn by 2029, worldwide (Statista, 2024). However, low-cost carriers (LCCs) are shifting the traditional perceptions of value and service among travelers by offering lower fares with fewer amenities. As traditional full-service airlines compete in this market environment, understanding the main drivers of customer satisfaction becomes critical to maintain competitiveness.

In this research, we analyze data from over 120,000+ airline passengers, including information on passenger demographics, flight characteristics (delays and distance), as well as service evaluations. Since this data lacks information on the specific airlines and ticket pricing, we will focus on analyzing the service-related factors that universally impact passenger satisfaction. This research aims to **determine the aspects of service that are important towards customer satisfaction.**

In our analysis, we initially explore the dataset to understand the variables and their distributions, as well as detect missing values. Data preparation procedures such as cleaning, normalization, and transformation are conducted before analysis. Modeling is performed by two different methods, decision tree classifier and logistic regression. Both models show a strong fit to the testing data. We further refine the model by utilizing boosting, assigning costs, as well dimensionality reduction. Results from both models converge well in identifying key drivers of satisfaction.

In summary, we find that Economy class passengers make up the majority of dissatisfied customers. However, a significant proportion of Business class travelers also report dissatisfaction. Using logistic regression, we explore the important aspects of service based on passenger class. The impact of each aspect of service were evaluated and ranked based on their magnitude (log-odds). We find that Business class and Economy class passengers had different drivers of satisfaction. For Business class passengers ease of online booking, delays, food and drink, as well as time convenience made up the dominant drivers for dissatisfaction. Whereas for Economy class passengers, in flight service, ease of online booking, and delay status were the dominant negative drivers. The full overview can be seen in Chapter 4.1. Recommendations for each driver are discussed in Chapter 4.2.

## Business Problem <a name="1.2"></a>

In this project, Group 17 serves as business consultants, with a focus on improving customer experience in the industry. Our responsibilities are from analyzing customer satisfaction statistics, developing measures to improve the airline's overall performance. Therefore, the business problem that we solve is "**How can airlines improve its strategy of services and activities to increase customer satisfaction, especially between segments and main service areas, to achieve stronger competitive advantages?**".

**The reason why we chose airline industry as our focus :**

-   Low-cost carriers (LCCs) are shifting the traditional perceptions of value and service among travelers by offering lower fares with fewer amenities.

-   Intense rivalry among airlines could be easily justified by customer satisfaction. Thus, higher customer satisfaction leads to increased client retention and overall business performance.

-   Concentrating on major market segments may provide a more general view of the aspects contributing to consumer satisfaction.

**Why focus on customer satisfaction?**

-   **Enhance customer loyalty**: improving on these key segments and service areas will enhance customer satisfaction that potentially increases customer loyalty rate.

-   **Competitive Advantage**: by addressing the service issue, the airline can stand out from the competitors.

## Setting Up the R Workplace <a name="1.3"></a>

Besides the Basic R Functions, this project uses multiple packages to enhance the R functionalities and improve the data analytics and visualization quality. This section explains all the packages and functions beyond the Basic R functionalities that we are using in this project in each step of the business analytics process from data cleaning to modeling.

------------------------------------------------------------------------

-   Data Cleaning and Preparation

    -   `dplyr` : used to filter rows, select columns, sort data, create new columns, and summarize data.

    -   `tidyr` : imported to be used to clean and tidy the data, such as reshaping the data from wide to long format.

-   Data Exploration

    -   `skimr` : designed to provide summary of statistical measures for each variables in data frames.

    -   `DataExplorer` : used to display comprehensive and convenient interface for data exploration and data visualization.

-   Data Transformation

-   Data Visualization

    -   `ggplot2` : a package that is useful to visualize data in a more complex and customizable way compared to the basic R plotting functions.

    -   `corrplot` : used to visualize correlation matrix between two variables in a more visually appealing and informative way.

-   Data Modeling

    -   `caret` : used for the process of creating predictive models, including data splitting, pre-processing, feature selection, and model tuning.

    -   `rsample` : a package to create resampling objects for both training and testing datasets.

    -   `c50` : used to create decision trees and rule-based models.

    -   `gmodel` : designed as a tool for fitting linear models.

    -   `ranger` : package for fast implementation of random forests with high cardinality data.

    -   `randomForest` : used to create a random forest model for the optimization of classification and decision tree.

    -   `pROC` : used to visualize the performance of classification models by plotting the ROC curve and further utilized to calculate the AUC (Area Under the Curve) score.
    
    -   `Information` : used to calculate the Information Value (IV) and Weight of Evidence (WOE) for each variable in the dataset.

-   Data Analysis

------------------------------------------------------------------------

Below we will find the commands to install the required libraries so that the code can run properly.The libraries only need to be imported only once, later we only need to call the necessary libraries when needed on the specific sections.

```{r, results='hide', warning=FALSE, message=FALSE}
# Importing Library

library(dplyr)
library(ggplot2)
library(skimr)
library(DataExplorer)
library(corrplot)
library(tidyr)
library(caret)
library(C50)
library(rsample)
library(gmodels)
library(ranger)
library(randomForest)
library(pROC)
library(Information)

```

## Dataset <a name="1.4"></a>

In this assignment, we use data set of [Airline Passenger Satisfaction Dataset](https://www.kaggle.com/datasets/mysarahmadbhat/airline-passenger-satisfaction) through Kaggle.The data consist of 24 variables from 129,880 passengers. Data cleaning and preparation are required to ensure the data is ready for analysis.

Firstly import all the data, csv file to be used in the data analysis process. The before using the data, it is important to understand all data type included in the data set. Subsequently, understanding statistic profile in the data set could help to prepare the data before using it.

```{r}
# Read Data
# Please note that the csv needs to be put into the same folders as this file for ensuring everything works properly
# The dataset should be downloaded from the Kaggle, or unzip from our submission in BrightSpace.
data <- read.csv("airline_passenger_satisfaction.csv")

#Understanding the data types
str(data)
```

```{r}
# Summary of Data statistics
summary(data)
```

### Data Dictionary

Based on Kaggle, here is the definition of each column in the dataset. This is useful for later stage to understand the data characteristics and how to use it.

| **Field Name** | **Description** |
|------------------------------------|------------------------------------|
| **ID** | Unique passenger identifier |
| **Gender** | Gender of the passenger (Female/Male) |
| **Age** | Age of the passenger |
| **Customer Type** | Type of airline customer (First-time/Returning) |
| **Type of Travel** | Purpose of the flight (Business/Personal) |
| **Class** | Travel class in the airplane for the passenger seat |
| **Flight Distance** | Flight distance in miles |
| **Departure Delay** | Flight departure delay in minutes |
| **Arrival Delay** | Flight arrival delay in minutes |
| **Departure and Arrival Time Convenience** | Satisfaction level with the convenience of the flight departure and arrival times from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Ease of Online Booking** | Satisfaction level with the online booking experience from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Check-in Service** | Satisfaction level with the check-in service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Online Boarding** | Satisfaction level with the online boarding experience from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Gate Location** | Satisfaction level with the gate location in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **On-board Service** | Satisfaction level with the on-boarding service in the airport from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Seat Comfort** | Satisfaction level with the comfort of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Leg Room Service** | Satisfaction level with the leg room of the airplane seat from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Cleanliness** | Satisfaction level with the cleanliness of the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Food and Drink** | Satisfaction level with the food and drinks on the airplane from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Service** | Satisfaction level with the in-flight service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Wifi Service** | Satisfaction level with the in-flight Wifi service from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **In-flight Entertainment** | Satisfaction level with the in-flight entertainment from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Baggage Handling** | Satisfaction level with the baggage handling from the airline from 1 (lowest) to 5 (highest) - 0 means "not applicable" |
| **Satisfaction** | Overall satisfaction level with the airline (Satisfied/Neutral or unsatisfied) |

## Workflow: Machine-Learning Task Translation <a name="1.5"></a>

The business analysis focuses on what key parameters from passengers' perspective that define their satisfactory level towards airline services. The analysis will be conducted by perfroming predictive modelling (regression analysis) as well as classification. Run code belows to see the workflow of the analysis.

```{r}
file_path <- "Workflow_Group17.png"
knitr::include_graphics(file_path)
```

## Reflection and Limitation <a name="1.6"></a>

In this research, we aim to predict **whether future customer would be satisfied given the details of service** by determining **the aspects of service that are important towards customer satisfaction.** From the dataset, we hypothesize that service characteristics such as arrival and departure delays, on-board service, and comfort play a significant role towards customer satisfaction.

Given that the dataset does not provide information on the specific airline or ticket pricing, we will focus to analyzing the service-related factors that universally impact passenger satisfaction. Understanding and preparing the the dataset will be crucial, as it may contain missing values or quality issues. Hence, we will dedicate the next section on data exploration and preparation. In doing so, we will first visualize the variables to get an understanding of the distributions.

# Data Exploration and Preparation <a name="2"></a>

In this chapter, we will use visual and descriptive data exploration methods to gain a deeper understanding of the distributions and the characteristics of variables in our dataset. We will begin by cleaning the data from any null variables, as well has handle outliers in the data. Later in this chapter, we will continue to explore each variable and interpret features. Finally, we will conduct initial plotting of the data to show how the features are correlated to each other.

## Data Cleaning and Preparation <a name="2.1"></a>

We will start the exploration phase by checking for missing data using the `is.na()` function.

```{r check_missing_data}

# Check for missing values
any(is.na(data))

# Compute percentage of missing values in each column
colMeans(is.na(data))
```

We find that null data only exists for 0.3% and only in the **Arrival Delay** column, therefore this would not significantly affecting the whole dataset. We'll fill the Arrival Delay null with the mean and median of Arrival Delay which is 0.

```{r eliminating_null_values}

# Replace NA values in Arrival.Delay with 0
data$Departure.Delay = as.numeric(data$Departure.Delay)
data$Arrival.Delay[is.na(data$Arrival.Delay)] <-0

# Check if there are any remaining NA values
any(is.na(data))
```

We succesfully replace the null values in the dataset. Let's check for numerical outliers using the `summary()` function, we will look for major differences between the median and the mean.

```{r summary_of_data}
# Extract summary statistics for numerical columns
numerical_summary <- data.frame(t(sapply(data[sapply(data, is.numeric)], summary)))

# Add a column to calculate the difference between mean and median
numerical_summary$Mean_Median_Difference <- abs(numerical_summary$Mean - numerical_summary$Median)

# Tabulate the output
knitr::kable(numerical_summary, caption = "Summary of Numerical Columns with Mean-Median Differences")

remove(numerical_summary)
```

We notice that there are large outliers in the Flight Distance, as well as in Arrival and Departure Delay. Let's see in the boxplot format :

**Flight Distance**

```{r}

# Boxplot for Flight Distance
ggplot(data, aes(x=Flight.Distance)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

**Departure Delay**

```{r}

# Boxplot for Departure Delay
ggplot(data, aes(x=Departure.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

**Arrival Delay**

```{r}

# Boxplot for Arrival Delay
ggplot(data, aes(x=Arrival.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

From the boxplots above, we can see that Flight Distance is slightly skewed while Arrival Delay and Departure Delay is heavily skewed. We will use median imputation, which is appropriate when the distribution of the data is skewed.

```{r median_imputation}
# Imputation of flight.distance variable
data$Flight.Distance[which(
  data$Flight.Distance > (quantile(data$Flight.Distance, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Flight.Distance, na.rm = TRUE))
)] <- median(data$Flight.Distance, na.rm = TRUE)

# Imputation of arrival.delay variable
data$Arrival.Delay[which(
  data$Arrival.Delay > (quantile(data$Arrival.Delay, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Arrival.Delay, na.rm = TRUE))
)] <- median(data$Arrival.Delay, na.rm = TRUE)

# Imputation of departure.delay variable
data$Departure.Delay[which(
  data$Departure.Delay > (quantile(data$Departure.Delay, 0.75, na.rm = TRUE) + 1.5 * IQR(data$Departure.Delay, na.rm = TRUE))
)] <- median(data$Departure.Delay, na.rm = TRUE)
```

Now let's check the distribution of the data after the imputation.

```{r histogram_of_imputed_variables}
plot_histogram(data[ , c('Flight.Distance','Arrival.Delay','Departure.Delay')], ncol=2,ggtheme = theme_minimal())
```

Below, we will also take a look again at the boxplots for Departure Delay and Arrival Delay. Note that before the imputation, the plots were skewed to the point that the boxes were almost no longer visible.

```{r}

# Boxplot for Departure Delay after imputation
ggplot(data, aes(x=Departure.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

```{r}

# Boxplot for Arrival Delay after imputation
ggplot(data, aes(x=Arrival.Delay)) +
geom_boxplot(fill = 2,alpha = 0.5,color = 1,outlier.colour = 2) +
theme_bw()
```

We can see that we have significantly normalized the skewness of the data. Thus, we have successfully eliminated all missing values and normalized all outliers in the data. Next, we will continue further into the data exploration phase.

## Data Exploration <a name="2.2"></a>

In this subchapter, we will begin by conducting descriptive univariate data analysis which is followed by association between numerical and categorical variables.

### Univariate Data Analysis

Now we visualize the distribution in each categorical column

```{r visualizing_categorical_variables}

# Bar chart will be used for categorical data
plot_bar(data, ncol=2, order_bar=TRUE, ggtheme = theme_minimal())
```

We can see that the gender distribution is equal. Furthermore, the majority type of travel is business which is also reflected in the customer type, where the majority is returning customers rather than first-time. Hence, increasing satisfaction of loyal customers will bring significant impact to the overall of airline customer satisfaction rate. Next, we will visualize the distribution in each numerical column. The ID column is left out as it is just a list of numbers.

```{r visualizing_numerical_variables}

# Numerical variables are visualized using histogram
plot_histogram(data[ , !names(data) %in% "ID"], ncol=2, ggtheme = theme_minimal())
```

From the visualizations above, it is not immediately clear which variables may be strongly influencing each other. Hence, we will continue the exploration by analyzing the associations between numerical variables.

### Association Between Two Numerical Variable

Now we will analyze the association between two numerical variables. We will use the `corrplot` package to visualize the correlation matrix of the numerical variables in the dataset. The correlation matrix will help us identify any strong correlations between the numerical variables.

::: scroll-box
```{r corrplot}
numeric_data <- data[sapply(data, is.numeric)]
numeric_data <- na.omit(numeric_data)

# Compute correlation matrix
cor_matrix <- cor(numeric_data)

# Print the correlation matrix
print(cor_matrix)
```
:::

```{r, fig.width=14, fig.height=12}
# Visualize the correlation matrix
options(repr.plot.width=14, repr.plot.height=12)
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         order = "hclust",
         tl.col = "black",
         tl.cex = 0.8,
         number.cex = 0.7,
         addCoef.col = "black",
         diag = FALSE)
```

Then we will filter the correlation matrix to find strong correlations (greater than 0.5 or less than -0.5) between numerical variables.

```{r}
# Discover strong correlation
cor_matrix_flat <- as.data.frame(as.table(cor_matrix))
filtered <- subset(cor_matrix_flat, abs(Freq) > 0.5 & Var1 != Var2)
filtered <- filtered[!duplicated(t(apply(filtered[, 1:2], 1, sort))), ]

print(filtered)
```

From the graph, we can see that there are some variables that have a strong correlation with each other, which is Seat Comfort, Leg Room Service, Cleanliness, Food & Drink, and In-flight Service exhibit strong positive correlations with each other. This suggests that customers who rate one aspect of the service highly tend to rate others positively as well. Furthermore, A high correlation between Departure Delay and Arrival Delay indicates that flights that depart late are also likely to arrive late. Other numerical factors such as Flight Distance may have weaker correlations with customer satisfaction, indicating that distance alone is not a major factor in satisfaction levels.

**In-flight Service vs On-board Service (in %)**

After we have analyzed the correlation between numerical variables, we will now analyze the association between categorical variables. We will use a heatmap to visualize the association between two categorical variables. The heatmap will show the percentage of each combination of ratings for the two variables.

```{r}
count_data <- as.data.frame(table(data$In.flight.Service, data$On.board.Service))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Service vs On-board Service (in %)",
       x = "In-flight Service Rating",
       y = "On-board Service Rating",
       fill = "Percentage") +
  theme_minimal()
```

The heatmap above shows a strong positive correlation between In-flight Service and On-board Service. This suggests that passengers who rate one aspect of the service highly also tend to rate the other positively. Darker shades indicate higher percentages, with the highest concentration (20.2%) at ( In-flight Service= 4, On-board Service = 4), suggesting a tendency for passengers to rate both aspects positively together.

**In-flight Service vs Food and Drink (in %)**

Now we will analyze the association between In-flight Service and Food and Drink with the same method.

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Food.and.Drink))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Food and Drink (in %)",
       x = "In-flight Entertainment Rating",
       y = "Food and Drink Rating",
       fill = "Percentage") +
  theme_minimal()
```

From the heatmap above, we can see that there is a strong positive correlation between In-flight Entertainment and Food and Drink. he most common rating combination is (4,4), accounting for 16.9% of responses, followed by (5,5) with 14.8%, indicating This suggests that passengers who rate the in-flight entertainment highly also tend to rate the food and drink positively.

**In-flight Entertainment vs Cleanliness (in %)**

Then, we will analyze the association between In-flight Entertainment and Cleanliness with the same method.

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Cleanliness))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Cleanliness (in %)",
       x = "In-flight Entertainment Rating",
       y = "Cleanliness Rating",
       fill = "Percentage") +
  theme_minimal()
```

We can see from the heatmap above that there is a positive correlation between In-flight Entertainment and Cleanliness with frequent rating pair is (4,4) at 17.9%, followed by (5,5) at 15.4%,. This suggests that passengers who rate the in-flight entertainment highly also tend to rate the cleanliness positively.

**In-flight Entertainment vs Seat Comfort (in %)**

Next,we analyze the In-flight Entertainment and Seat Comfort with the same method.

```{r}
count_data <- as.data.frame(table(data$In.flight.Entertainment, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-flight Entertainment vs Seat Comfort (in %)",
       x = "In-flight Entertainment Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

This heatmap depicts a positive correlation between In-flight Entertainment and Seat Comfort. The most common rating combination is (4,4) at 17.9%, followed by (5,5) at 15.1%. This suggests that passengers who rate the in-flight entertainment highly also tend to rate the seat comfort positively.

**Food and Drink vs Cleanliness (in %)**

```{r}
count_data <- as.data.frame(table(data$Food.and.Drink, data$Cleanliness))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Food and Drink vs Cleanliness (in %)",
       x = "Food and Drink Rating",
       y = "Cleanliness Rating",
       fill = "Percentage") +
  theme_minimal()
```

**Food and Drink vs Seat Comfront (in %)**

Now we will analyze the association between Food and Drink and Seat Comfort with the same method.

```{r}
count_data <- as.data.frame(table(data$Food.and.Drink, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Food and Drink vs Seat Comfort (in %)",
       x = "Food and Drink Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

This heatmap illustrates a positive correlation between Food and Drink and Seat Comfort. The most common rating combination is (4,4) at 16%, followed by (5,5) at 13.8%. This suggests that passengers who rate the food and drink highly also tend to rate the seat comfort positively.

**Cleanliness vs Seat Comfort (in %)**

Afterwards, we will analyze the association between Cleanliness and Seat Comfort.

```{r}
count_data <- as.data.frame(table(data$Cleanliness, data$Seat.Comfort))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "Cleanliness vs Seat Comfort (in %)",
       x = "Cleanliness Rating",
       y = "Seat Comfort Rating",
       fill = "Percentage") +
  theme_minimal()
```

This heatmap shows a positive correlation between Cleanliness and Seat Comfort. The most common rating combination is (4,4) at 17.9%, followed by (5,5) at 15.1%. This suggests that passengers who rate the cleanliness highly also tend to rate the seat comfort positively.

**In-Flight Wifi Service vs Ease of Online Booking (in %)**

Next, we will analyze the association between In-flight Wifi Service and Ease of Online Booking.

```{r}
count_data <- as.data.frame(table(data$In.flight.Wifi.Service, data$Ease.of.Online.Booking))
count_data$Percent <- round(100 * count_data$Freq / sum(count_data$Freq), 1)

# Plot heatmap
ggplot(count_data, aes(x = Var1, y = Var2, fill = Percent)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black", size = 3) +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(title = "In-Flight Wifi Service vs Ease of Online Booking (in %)",
       x = "In-Flight Wifi Service Rating",
       y = "Ease of Online Booking Rating",
       fill = "Percentage") +
  theme_minimal()
```

This heatmap shows a positive correlation between In-flight Wifi Service and Ease of Online Booking. The most common rating pair is (2,3) at 18.1%, closely followed by (1,2) at 18% and (1,1) at 12.5%, suggesting that many passengers who rated WiFi service poorly also found online booking less convenient. Interestingly, higher ratings for online booking (4-5) are associated with mixed ratings for WiFi, with 8.1% rating both as (5,5). This pattern implies that while passengers may find booking easy, WiFi service still receives moderate to low ratings.

**Flight Distance vs Departure Delay**

Next, we will analyze the association between Flight Distance and Departure Delay using scatter plot.

```{r}
ggplot(data, aes(x = Flight.Distance, y = Departure.Delay)) +
  geom_point(alpha = 0.2, color = "steelblue", size = 1.5) +
  labs(title = "Flight Distance vs Departure Delay",
       x = "Flight Distance",
       y = "Departure Delay") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
  theme_minimal()
```

From the scatter plot above, we can see that there is a slight positive correlation between Flight Distance and Departure Delay. The red line likely represents a trend or regression line, showing the general pattern in the data. The data appears to be highly discrete, with many flights experiencing common delay times (e.g., 0, 5, 10, 15 minutes). The trend line suggests that departure delays slightly increase as flight distance increases, but the relationship is not strong. There is significant variability in departure delays across all flight distances, meaning other factors likely influence delays beyond just distance.

**Flight Distance vs Arrival Delay**

Subsequently, we will analyze the association between Flight Distance and Arrival Delay using scatter plot.

```{r}
ggplot(data, aes(x = Flight.Distance, y = Arrival.Delay)) +
  geom_point(alpha = 0.2, color = "steelblue", size = 1.5) +
  labs(title = "Flight Distance vs Arrival Delay",
       x = "Flight Distance",
       y = "Arrival Delay") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linewidth = 1) +
  theme_minimal()
```

The scatter plot above shows a slight positive correlation between Flight Distance and Arrival Delay. The red line, likely a trend or regression line, suggests that the average arrival delay remains relatively low regardless of flight distance. While there is some variability, the data does not indicate a strong correlation between flight distance and arrival delay.

### Association Between Numerical and Categorical Variable

Now, we will analyze the association between numerical and categorical variables. We will use boxplots to visualize the distribution of numerical variables across different categories.

```{r}
data_subset <- data %>%
  select(Satisfaction, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Satisfaction, y = Value, fill = Satisfaction)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Satisfaction",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

The boxplots compare numeric variables (Age, Arrival Delay, Departure Delay, and Flight Distance) by satisfaction levels (Neutral or Dissatisfied vs. Satisfied). Satisfied customers tend to have higher ages and longer flight distances, as indicated by the higher medians and broader distributions for these groups. Conversely, Arrival and Departure Delays show minimal differences between satisfaction groups, with both distributions being highly skewed and overlapping significantly. This suggests that delays may not strongly influence satisfaction compared to age and flight distance.

```{r}
data_subset <- data %>%
  select(Satisfaction, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Satisfaction, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

The histograms display the distribution of numeric variables grouped by customer satisfaction levels which is Neutral or Dissatisfied (pink) and Satisfied (blue) across various aspects of airline services. In general, higher ratings (closer to 5) for services such as Check-in Service, Ease of Online Booking, Leg Room Service, Online Boarding, and Gate Location correspond more strongly with satisfied customers. Conversely, dissatisfaction is more prevalent at lower ratings (closer to 0 or 1). For Flight Distance, satisfied customers are slightly more common in longer flights. Similarly, variables like Baggage Handling, Cleanliness, Food and Drink, In-flight Entertainment, In-flight Service, In-flight Wi-Fi Service, and Seat Comfort show a similar trend where higher ratings indicate greater satisfaction, while lower ratings are associated with dissatisfaction. These distributions suggest that satisfaction is closely tied to the quality of services provided across these dimensions.

```{r}
data_subset <- data %>%
  select(Class, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Class, y = Value, fill = Class)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Class",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

The boxplots compare Age, Arrival Delay, Departure Delay, and Flight Distance across Business, Economy, and Economy Plus classes. Business class passengers are generally older, with a higher median age compared to Economy and Economy Plus, which show similar distributions. Arrival and departure delays have medians near zero across all classes, indicating most flights are on time, but all classes exhibit outliers representing extreme delays. Flight distance is notably higher for Business class, with a greater median and variability compared to the shorter and more consistent distances in Economy and Economy Plus. This suggests Business class is preferred for longer flights and by older passengers, while delays are consistent across all classes.

```{r}
data_subset <- data %>%
  select(Class, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Class)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Class",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Class, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Class, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Class)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Class",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

The histograms compare numeric variables across flight classes (Business, Economy, and Economy Plus) to assess customer satisfaction. Business class consistently receives higher ratings across most variables, such as Check-in Service, Leg Room Service, Online Boarding, On-board Service, Cleanliness, Food and Drink, In-flight Entertainment, In-flight Wifi Service, and Seat Comfort, highlighting its superior service quality and comfort. Economy Plus generally performs better than Economy but falls short of Business class ratings. Economy class exhibits greater variability in ratings and is often skewed toward lower scores for services like legroom, onboard service, food quality, and seat comfort. Flight Distance shows Business class dominating longer flights, while Economy handles shorter distances. Variables like Gate Location and Ease of Online Booking show relatively even distributions across classes. Overall, the data emphasizes the premium experience associated with Business class compared to Economy and Economy Plus.

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Type.of.Travel, y = Value, fill = Type.of.Travel)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Type of Travel",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

The boxplots illustrate the distribution of four numeric variables,Age, Arrival Delay, Departure Delay, and Flight Distance grouped by the type of travel (Business or Personal). Business travelers tend to be slightly older on average than personal travelers, with a narrower age range. Both groups experience similar patterns in arrival and departure delays, though delays for personal travel show more variability and outliers. Flight distance varies significantly between the two groups, with business travelers generally covering shorter distances compared to personal travelers, which exhibit a wider range and more outliers.

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Type.of.Travel)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Type of Travel",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Type.of.Travel, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Type.of.Travel, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Type.of.Travel)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Type of Travel",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

The histograms illustrate, Across most variables, Business travelers dominate the counts, particularly for higher ratings (4 and 5), indicating greater satisfaction. In contrast, Personal travelers are fewer and show a more even distribution across lower and higher ratings. Flight Distance shows a distinct peak for shorter distances, with Business travelers being more frequent. Overall, Business travelers tend to rate services higher than Personal travelers.

```{r}
data_subset <- data %>%
  select(Gender, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Gender, y = Value, fill = Gender)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Gender",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

The boxplots compare four numeric variables of Age, Arrival Delay, Departure Delay, and Flight Distance by gender (Female and Male). For Age, males have a slightly higher median than females, with similar interquartile ranges (IQRs) and no visible outliers. Arrival Delay and Departure Delay show comparable distributions for both genders, with medians near zero, narrow IQRs, and numerous outliers indicating extreme delays. Flight Distance reveals a slightly higher median and wider variability among males compared to females, with some outliers in both groups representing unusually long flights. Overall, the distributions of delays are similar across genders, while Flight Distance shows more variability for males.

```{r}
data_subset <- data %>%
  select(Gender, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Gender)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Gender",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Gender, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Gender, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Gender)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Gender",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

The histograms illustrate the distribution of ratings for various airline service attributes (Baggage Handling, Cleanliness, Food and Drink, In-flight Entertainment, In-flight Service, In-flight WiFi Service, and Seat Comfort) segmented by gender (Female and Male). Across most categories, ratings of 4 and 5 are predominant, indicating overall positive feedback. Both genders exhibit similar patterns in their evaluations, with slight variations in proportions. For example, females appear to rate slightly higher in categories like In-flight Entertainment and Cleanliness compared to males. Conversely, males seem to have a marginally higher count in lower ratings (1–3) for some categories like Seat Comfort and In-flight WiFi Service. This suggests gender-based nuances in satisfaction levels but overall consistency in favoring higher ratings across services

```{r}
data_subset <- data %>%
  select(Customer.Type, Age, Flight.Distance, Departure.Delay,
         Arrival.Delay)

# Convert to long format for faceting
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Plot boxplots with facets
ggplot(data_long, aes(x = Customer.Type, y = Value, fill = Customer.Type)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_minimal() +
  labs(title = "Boxplots of Numeric Variables by Customer Type",
       x = NULL, y = NULL) +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "none"
  )
```

The boxplots compare numeric variables (Age, Arrival Delay, Departure Delay, and Flight Distance) between two customer types: First-time and Returning. For Age, returning customers tend to be older, with a higher median and wider range compared to first-time customers. Arrival Delay and Departure Delay show similar distributions for both groups, with most values concentrated around zero and a significant number of outliers indicating occasional long delays. For Flight Distance, returning customers generally travel longer distances, as evidenced by a higher median and greater variability compared to first-time customers. Overall, returning customers exhibit distinct patterns in age and flight distance, while delays are similarly distributed across both groups.

```{r}
data_subset <- data %>%
  select(Customer.Type, Ease.of.Online.Booking, Check.in.Service, Online.Boarding,
         Gate.Location, On.board.Service, Leg.Room.Service, Flight.Distance)

# Convert to long format
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Customer.Type)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Customer Type",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

```{r}
data_subset <- data %>%
  select(Customer.Type, Seat.Comfort, In.flight.Service, Baggage.Handling,
         In.flight.Entertainment, In.flight.Wifi.Service, Food.and.Drink, Cleanliness)

# Convert to long format
data_long <- pivot_longer(data_subset, -Customer.Type, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Customer.Type)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables by Customer Type",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )
```

The two sets of histograms display the distribution of numeric variables related to customer satisfaction across two customer types: "First-time" and "Returning." Each variable, such as "Check.in.Service," "Flight.Distance," "Leg.Room.Service," and others, is rated on a scale (likely 0–5), and the counts for each rating are shown separately for the two customer types, with "Returning" customers represented in teal and "First-time" customers in pink. Across most variables, returning customers consistently dominate the counts, suggesting they are more frequent or engaged with the airline. Ratings of 4 and 5 tend to have higher counts for both groups in service-related categories like "Ease.of.Online.Booking," "Leg.Room.Service," and "Seat.Comfort," indicating general satisfaction. However, some variables like "Gate.Location" and "In.flight.Wifi.Service" show substantial variability, suggesting room for improvement in these areas. Overall, the histograms provide insights into customer preferences and satisfaction trends based on experience type.

### Association Between Two Categorical Variable

Recalling the Univariate Data Analysis, we will use five categories or variables that we deemed important as the base for further analysis, specifically to see the association between two categorical variables. These variables are distribution of gender, customer type, type of travel, class, and the disparity of numbers between satisfacted and unsatisfied passengers.

```{r}
plot_bar(data,ncol=2)
```

We will use the plot_bar function to visualize the association between these variables. For each of the following five sections, we take out one of the category and plot the association between each of the remaining four categories with the missing category. We start with creating the association between customer's satisfaction with the other four variables.

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```

From the univariate data distribution, we already know the numbers of satisfied and unsatisfied passengers is not equal with noticeable difference, with the majority of passengers being unsatisfied. So it is not surprising that regardless of the association between passengers' satisfaction with other categories, most graphs are showing the skew towards the neutral or dissatisfied category.

However, it could be noticed that the distribution of satisfaction level across both genders are relatively equal, while the first time flyers are more likely to be unsatisfied rather than the frequent travelers. Business travelers also tend to be more satisfied than the economy class passengers. This is also the case for the business class passengers. We will see in the other section that passengers with business intention are more likely to fly on business class, so the correlation is consistent in the satisfaction level as well. This is understandable, since airlines with business class are more likely to have extra services and better quality of basic services, which means that it is more likely to satisfy the passengers. Based on this information, the focus on improving the satisfaction rate of passengers should be on the economy class passengers with leisure intention, especially those who are first time flyers.

```{r}
plot_bar(data,by='Gender',ncol=2)
```

In this section, we show the distribution of genders in four other categories, which is relatively balanced across all graphs. Hence, we can conclude that genders are not the deciding factor that we need to analyze further in the data analysis section, meaning that it is not important enough as a consideration point to improve the satisfaction rate.

```{r}
plot_bar(data,by='Customer.Type',ncol=2)
```

From the customer type distribution and association graphs, it can be noticed that the majority of first time passengers are flying economy class, and also more likely to have the less-favorable first imnpression towards their flying experience, which can be seen in its relation with satisfaction level.

```{r}
plot_bar(data,by='Type.of.Travel',ncol=2)
```

As we can see above, the majority of business travelers fly on the business class as well, and they are also more likely to be satisfied with their flying experience. This is consistent with our observation and discussion on the previous section. On the opposite, travelers with personal reason, such as leisure, are more likely to fly on both of the economy classes and also less likely to be satisfied with their flying experience.

```{r}
plot_bar(data,by='Class',ncol=2)
```

The plot above shows the distribution of passengers divided by the class of service they fly in. Again, we can see that it the business-oriented travelers are also dominated by business class flyers, and they (business class passengers) are also more likely to be satisfied with their flying experience. This is consistent with our observation and discussion on the previous section. On the opposite, both economy and economy plus class passengers are less likely to be satisfied with their flying experience and more likely to be first time flyers.

## Feature Engineering <a name="2.3"></a>

We noticed that the age distribution have high variation, therefore we try to create the age_group column to divide the age group into four categories: children, youth, adult, and senior. This categorization is based on most of the division range used in a lot of transportation services across the world. The age group is defined and plotted in R as follows:

```{r}
data = data %>% mutate(Age_Range = if_else(Age <= 15,"Children",
if_else(16 <= Age & Age <= 25,"Youth",
if_else(26 < Age & Age <= 60,"Adult","Senior"))))
```

```{r}
plot_bar(data$Age_Range,ncol=2,order_bar=T,ggtheme = theme_minimal(), title = "Age Group")
```

Now we can interpret from the age group that the largest segment of our customer is belong to adult age group. This is understandable since the adult group has the widest range of age compared to other age groups.

Another data we can group is the flight distance which currently right skewed. We will create new variables of flight distance by grouping them into three categories: short-haul, medium-haul, and long-haul. The division of distance is based on the limit defined by David W. Wragg in his book A Dictionary of Aviation:

**Short-haul Flights**: 0 - 900 miles (0 - 1,450 km)

**Medium-haul Flights**: 900 - 2,200 miles (1,450 - 3,540 km)

**Long-haul Flights**: 2,200 miles (3,540 km) and above

```{r}
data = data %>% mutate(Distance_Group = if_else(Flight.Distance < 900,"Short-haul",
if_else(900 <= Flight.Distance & Flight.Distance <= 2200,"Medium-haul","Long-haul")))
```

```{r}

plot_bar(data$Distance_Group, ncol = 2, order_bar = TRUE, ggtheme = theme_minimal(), title = "Flight Distance Group")

```

We can divide which customer experiencing delay or not based on departure delay and arrival delay. United States' Federal Aviation Administration (FAA) defines a flight as delayed if it is 15 minutes or more behind schedule. We can use the following code to create the new variables to show the comparison of the numbers of flight with departure delay and arrival delay.

```{r}
data$departure.delay.status = ifelse(data$Departure.Delay > 15, "Delayed", "On Time")
data$arrival.delay.status = ifelse(data$Arrival.Delay > 15, "Delayed", "On Time")

plot_bar(data$departure.delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal(), title = "Departure Delay Status")
plot_bar(data$arrival.delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal(), title = "Arrival Delay Status")
```

Because both departure and arrival delay status are equally important, we can combine them into one variable called **delay.status** with OR operator is used to display the total number of delayed flights.

```{r}
data$delay.status = ifelse(data$Departure.Delay > 15 | data$Arrival.Delay > 15, "Delayed", "On-time")

plot_bar(data$delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal(), title = "Delay Status")
```

Next, We group the features into category of services that could be improved by the airline:

-   **Flight General Service:** On-board Service, Food and Drink, In-flight Service

-   **Flight Entertainment:** In Flight Wifi Service, In Flight Entertainment

-   **Pre-flight Service:** Ease of Online Booking, Check-in Service, Online Boarding

-   **Comfortability:** Seat Comfort, Leg Room Service, Cleanliness

```{r}
# Create features group with maximum score of 5 
data$Flight.general.service = rowSums(data[,c('On.board.Service','Food.and.Drink','In.flight.Service')],na.rm = TRUE)/3
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)/2
data$Pre.flight.service = rowSums(data[,c('Ease.of.Online.Booking','Check.in.Service','Online.Boarding')],na.rm = TRUE)/3
data$Comfortability = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)/3

plot_histogram(data[ , c("Flight.general.service", "Flight.Entertainment", "Pre.flight.service", "Comfortability")], ncol=2,ggtheme = theme_minimal())
```

Combining several category of services into one variable is useful to see the overall satisfaction of the airline services in numerous area of improvement. For example, as we can see above that in general the passengers found that the flights they are taking are relatively comfortable. However, the pre-flight service and flight entertainment category received mixed reviews and can be considered as an improvement area for the airline, where further analysis could be done to see which of the element of those services are the most dissatisfying for the passengers.

## Data Transformation <a name="2.4"></a>

Data transformation plays a key role to make sure all data is in the same scale and distribution. We will start to take a look on our data distribution.

```{r}

# Set default colors for histogram bars
update_geom_defaults("bar", list(fill = "steelblue", color = NA))

# Generate histogram with updated colors
plot_histogram(data, 
               ncol = 2, 
               ggtheme = theme_minimal() + theme(strip.background = element_blank(),
                                                 strip.text = element_text(color = "black")))
```

This part is to undertand the distribution of data before it is used. As we can see in the histogram, the data is not normally distributed. Therefore, we need to normalize the data. There are several concerning variables in the dataset that are part of main focus in the data transformation process.

<ul>

<li>**arrival.delays** and **departure.delays** have highly-right skewed distribution where most values are concentrated to zero.</li>

<li>**flight.distance** has a right-skewed distribution concentrated between 500 and 1500, with a long tail up to 4000.</li>

<li>**age** has a right-skewed distribution concentrated between 20 and 50, with a long tail up to 80.But this variable is less concerning compared to the previous ones.</li>

</ul>

In the following step we will normalize the data by applying log transformation to the concerning variables.

```{r}
# Create new variables for departure and arrival delay duration
data$Departure.Delay.Duration = ifelse(data$Departure.Delay > 0, data$Departure.Delay, NA)
data$Arrival.Delay.Duration = ifelse(data$Arrival.Delay > 0, data$Arrival.Delay, NA)

# Apply log transformation to the variables
data$Flight.Distance.log = log(1 + data$Flight.Distance)
data$Age.log = log(1 + data$Age)
data$Departure.Delay.Duration.log = log(1 + data$Departure.Delay.Duration)
data$Arrival.Delay.Duration.log = log(1 + data$Arrival.Delay.Duration)

# Plot histograms of the log-transformed variables
plot_histogram(data[, c("Flight.Distance.log", "Age.log", "Departure.Delay.Duration.log", "Arrival.Delay.Duration.log")], ncol=2,ggtheme = theme_minimal())
```

Data transformation is performed with visualization task to see the distribution of the data after the transformation. This aims to normalize skewed distribution.

<ul>

<li>**arrival.delays** and **departure.delays** are transformed into log scale to address the skewed issues.</li>

<li>**flight.distance** and **age** are also transformed into log scale to address the skewed issues.</li>

</ul>

After log-transormation, normalize the distribution is then performed as follow.

```{r}
# Normalize the log-transformed variables
data$Flight.Distance.log.z = scale(data$Flight.Distance.log)
data$Age.log.z = scale(data$Age.log)
data$Departure.Delay.Duration.log.z = scale(data$Departure.Delay.Duration.log)
data$Arrival.Delay.Duration.log.z = scale(data$Arrival.Delay.Duration.log)

# Plot histograms of the normalized variables
plot_histogram(data[, c("Flight.Distance.log.z", "Age.log.z", "Departure.Delay.Duration.log.z", "Arrival.Delay.Duration.log.z")], ncol=2,ggtheme = theme_minimal())
```

As we can see in the histogram, the data is now normalized with 0 as the center. This is important to ensure that the data is on the same scale and distribution. This helps the model learn the data better and make better predictions.

Here are the reasons why performing normalization is important before using the data:

<ol>

<li><strong>Comparison Across Variables</strong><br> Normalization ensures the data is on the same scale. This helps in making the model comparable and supports easier further analysis.</li>

<li><strong>Improved Model Performance</strong><br> Many algorithms perform better when input variables are standardized, as it prevents variables with larger ranges from dominating the analysis.</li>

<li><strong>Normalization Beyond Log Transformation</strong><br> While log transformation reduces skewness and normalizes distributions, scaling ensures uniformity across variables, enabling fair use in statistical techniques.</li>

</ol>

## Initial Plotting <a name="2.5"></a>

Our target is the satisfaction rate of the airlines services, therefore we need to identify how each variable are correlated to the satisfaction. This is important to understand how each variable's behavior toward main objective of the analysis. Following is the behavior of each variables towards satisfaction rate.

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```

According to the graphic above, we can see that:

<ul>

<li>**Gender**: Both male and female customers have a higher proportion of "Neutral or Dissatisfied" responses compared to "Satisfied" ones. This indicates there is a room of improvement to increase satisfied number</li>

<li>**Customer Type**: Returning customers exhibit a higher proportion of "Satisfied" responses than first-time customers, who seem mostly "Neutral or Dissatisfied". This might indicate that familiarity with the company positively influences satisfaction.</li>

<li>**Type of Travel**: Business travelers show a higher proportion of "Satisfied" responses compared to personal travelers.</li>

<li>**Class**: Business class has the highest proportion of "Satisfied" responses, whereas Economy and Economy Plus lean more towards "Neutral or Dissatisfied." This indicates higher satisfaction among premium customers.</li>

<li>**Age Range**: Adult and youth customers show higher satisfaction levels compared to senior and children, who lean towards dissatisfaction.</li>

<li>**Distance Group**: Long Haul seem to lead to higher satisfaction levels compared to short haul and medium haul, which could indicate that services during these trips meet or exceed customer expectations.</li>

<li>**Delay Status**: The behavior of satisfaction between on-time and delayed status is not significantly different</li>

</ul>

```{r}
# Create a subset of the data with grouped airline services
data_subset <- data %>%
  select(Satisfaction, Flight.general.service, Flight.Entertainment, Pre.flight.service, Comfortability)

# Convert to long format
data_long <- pivot_longer(data_subset, -Satisfaction, names_to = "Variable", values_to = "Value")

# Create faceted histogram
ggplot(data_long, aes(x = Value, fill = Satisfaction)) +
  geom_histogram(position = "stack", alpha = 0.5, bins = 30) +
  facet_wrap(~ Variable, scales = "free", ncol = 2) +
  theme_minimal() +
  labs(title = "Histograms of New Numeric Variables by Satisfaction",
       x = NULL, y = "Count") +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom"
  )

```

Graphic above is telling the satisfaction distribution based on services category such as comfortability, flight entertainment, in flight service, and pre flight service. We can see that the distribution of satisfaction is not significantly different between each category.

```{r, message=FALSE, warning=FALSE}

# Percentage of satisfaction by Class
data %>% group_by(Class, Satisfaction) %>% summarize(num=n()) %>% mutate(percentage=round(num*100/sum(num),2))

```

Based on the data above, **69.4%** of Business passengers showing satisfaction, while only **18.8%** of Economy passengers showing satisfaction and only **24.6%** of Economy Plus passengers are satisfied. This indicates that the airline services are more satisfying for business class passengers compared to economy class passengers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly in Economy and Economy Plus, to achieve better satisfaction rates.

```{r, message=FALSE, warning=FALSE}

# Percentage of satisfaction by Type of Travel
data %>% group_by(Type.of.Travel, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```

Personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only **10%** from it. This is much lower compared to the business type customers with **58.4%** satisfaction rate. Personal travel satisfaction levels indicate areas for improvement in service offerings or customer experience. To enhance customer satisfaction among various traveler types, it is crucial to refine customer segmentation strategies. This involves identifying and analyzing distinct traveler segments to understand their unique satisfaction triggers. By doing so, strategies to improve satisfaction rate on each segments can be effectively implemented.

```{r, message=FALSE, warning=FALSE}

# Percentage of satisfaction by Customer Type
data %>% group_by(Customer.Type, Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```

**First-time Customers**: In this customer segment, ration between dissatisfaction customers compared to satisfied customers is 5:1, indicating a significant gap in satisfaction. The data shows around 24% of first-time customers are satisfied, while around 76% are neutral or dissatisfied. This indicates that first-time customers are not satisfied with the airline services, which is a concern for the airline as it indicates a risk of losing first-time customers.<br>

**Returning Customers**: A majority, around 52%, are satisfied, while around 48% are neutral or dissatisfied. This indicates that returning customers are more satisfied with the airline services compared to first-time customers. This data highlights the need for airlines to focus on enhancing the customer experience, particularly for first-time customers, to achieve better satisfaction rates.

```{r, message=FALSE, warning=FALSE}

# Percentage of satisfaction by Class for First-time Customers
data %>% filter(Customer.Type == 'First-time') %>% group_by(Class ,Satisfaction) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))

```

The table shows significant differences in first-timer passenger satisfaction across flight classes. Business Class shows the highest satisfaction, with nearly 40% satisfied, though more than half remain neutral or dissatisfied. Economy Class fares worse, with 85% of passengers are not satisfied, and Economy Plus performs the poorest, with 92% dissatisfaction. Enhancing services for Economy and Economy Plus could greatly improve overall customer satisfaction.<br> <br> In conclusion, mostly first-timer passengers feeling dissatisfied on each classes with notably higher in economy and economy plus class. This is a concern for the airline, as it indicates a need to improve services for first-time customers, to retain them and increase satisfaction rates. Leaving a good impression for first timer consumers is important to ensure they will come back and use the airline services again in the future. This is also important to ensure that the airline can maintain its market share and profitability in the long run.

```{r}
data %>% group_by(Distance_Group) %>%
summarize(num=n()) %>%
mutate(percentage=round(num*100/sum(num),2))
```

Previously in the barplot satisfaction, we can notice that as the distance lower the rate of dissatisfied customer is increasing. **short-haul trip is having the highest dissatisfaction rate**. From the data above, this condition could be a big concern since the majority of the airline customer is taking short-haul trip (more than 60%). This is a concern for the airline, as it indicates a need to improve services for short-distance travelers, to retain them and increase satisfaction rates. Failing to retain loyalty from the majority of customers (short-haul customers) could lead to a significant loss in market share and profitability. <br> <br> Next we will try to analyze the service factors related to the customer satisfaction focusing on low satisfaction profile, starting with customer type : <br><br>
<span style="color: #1E90FF; font-size: 16px;">**CUSTOMER TYPE: FIRST-TIME**</span><br>


```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Customer.Type == 'First-time') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, 
                                                   Departure.Delay, Arrival.Delay.Duration, 
                                                   Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, 
                                                   Departure.Delay.Duration, 
                                                   Departure.Delay.Duration.log, 
                                                   Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, 
                                                   Age.log, Age.log.z))

# Remove NA values (non-numeric columns)
mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)  
mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#CFCFFF") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

As discussed above, first-timer group showed a significant dissatsifaction rate, thus through this graph, we can analyze which triggers the dissatisfaction. The graph above shows the mean of each service category for first-timer customers. The mean values are sorted in descending order, with the highest mean value on top. The mean values are calculated from the satisfaction rating of each service category. <br>

We can identify that from the first time customer, the services that dissapointing them most are **Departure and Arrival Time Convenience, Ease of Online Booking, and Wifi Service**.While **In flight service and comfortability** having the highest satisfaction value. <br>

This data can guide strategic improvements to enhance the first-time customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score.<br><br>
<span style="color: #1E90FF; font-size: 16px;">**TYPE OF TRAVEL: PERSONAL**</span><br>

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Type.of.Travel == 'Personal') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, 
                                                   Departure.Delay, Arrival.Delay.Duration, 
                                                   Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, 
                                                   Departure.Delay.Duration, 
                                                   Departure.Delay.Duration.log, 
                                                   Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, 
                                                   Age.log, Age.log.z))

# Remove NA values (non-numeric columns)
mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)  
mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#008080") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")

```

In the previous section, we can see that personal type of travel passenger is having a third of airline total customer with the number of satisfied customer is only **10%** from it. Thus, the graph above becomes important to understand which factors are contributing to the low satisfaction rate. The graph above shows the mean of each service category for personal type of travel customers. The mean values are sorted in descending order, with the highest mean value on top. The mean values are calculated from the satisfaction rating of each service category. <br>

We can identify that from the Business Travel type of travel customers, the services that dissapointing them most in **Ease of online booking, In flight wifi service, online boarding**.These areas could be prioritized for enhancement to improve overall satisfaction for business travelers. The top satisfaction factors are **in flight service and Comfortability**.<br>

As you can see, personal type of travel customer satisfaction data is showing similar behaviour with the first-timer passengers. Thus, improving low-rated areas like WiFi availability and online booking ease may help elevate the overall customer experience for this segment. <br><br>
<span style="color: #1E90FF; font-size: 16px;">**CLASS: ECONOMY**</span><br>

```{r}

# Compute mean for each continuous column
filtered_data <- data %>% filter(Class == 'Economy') 
filtered_data <- subset(filtered_data, select = -c(ID, Flight.Distance, Age, Arrival.Delay, 
                                                   Departure.Delay, Arrival.Delay.Duration, 
                                                   Arrival.Delay.Duration.log,
                                                   Arrival.Delay.Duration.log.z, 
                                                   Departure.Delay.Duration, 
                                                   Departure.Delay.Duration.log, 
                                                   Departure.Delay.Duration.log.z,
                                                   Flight.Distance.log, Flight.Distance.log.z, 
                                                   Age.log, Age.log.z))

# Remove NA values (non-numeric columns)
mean_values <- sapply(filtered_data, function(x) if(is.numeric(x)) mean(x, na.rm = TRUE) else NA)
mean_values <- na.omit(mean_values)  
mean_df <- data.frame(Variable = names(mean_values), Mean = mean_values)

# Sort in descending order
mean_df <- mean_df[order(mean_df$Mean, decreasing = TRUE), ]

# Plot bar chart with highest value on top and number labels
ggplot(mean_df, aes(x = reorder(Variable, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill = "#DC143C") +
  geom_text(aes(label = round(Mean, 2)), hjust = -0.2) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Mean of Continuous Variables", x = "Variable", y = "Mean Value")
```

From the previous discussion, we identify that customers in economy class having high number of dissatisfaction. The graph above is helping to identify which parameters should have more attention to be improved for higher customers satisfaction rate. <br> The graph above is satisfaction data from Economy class customers. The biggest contributors to low satisfaction rate are **ease of online booking, wifi service, and online boarding**. While **In flight service and Comfortability** have the highest satisfaction rate among the Economy class customers. This data can guide strategic improvements to enhance the Economy class customer experience and increase satisfaction rates by focusing the improvement on those concerning satisfaction factors with the low score. <br><br>
The mean values of the continuous variables, as displayed above, provide a preliminary understanding of the dataset. However, these raw values are not directly suitable for decision-making or interpretation in their current form. Further refinement will be conducted through advanced modeling techniques, specifically classification and logistic regression, to derive actionable insights and meaningful patterns.<br>


### Summary of Initial Plotting

<ol>

<li>43% of total passengers have liked the airline service. The majority of customers who are satisfied are from business class</li>

<li>57% of total passengers are not satisfied. The majority of them are from Economy class</li>

<li>Only 18% of people have travelled for first time in our airline service. Out of those customers, nearly 76% of customers are not satisfied with our service which could create a chance in future of not returning of customers to our service again. When passengers are coming for first time, most of them have chosen the economy class</li>

<li>Out of the customers who opted for business class, nearly 69% of customers are satisfied with the business class. Out of the customers who opted for economy class, nearly 81% of them are not satisfied with it. Customers who opted for economy plus class, nearly 75% of them are not satisfied with it.</li>

<li>Whether people are coming for first time or not and also irrespective of their travel type, they are mostly not satisfied with the Economy and Economy Plus classes. There is strong need for the improvement of Economy class since 45% of total passengers are from economy class. Moreover, 60% of passengers travel less than 10k km and most of them prefer economy class.</li>

<li>Passengers are not so satisfied with online booking, Online Boarding and the Wifi service mostly especially those who take economy class.</li>

<li>82% of total passengers are returning to the airline service. Out of those 105773 passengers, 50% of passengers are from business class and 42% of passengers are from economy class. Customer who are returning to our service are aged above 30 and make business travel and travel less than 1000 km mostly</li>

<li>Departure and arrival time convenience is considered big contributors to the customers dissatisfaction especially for first-timer passengers and personal travel type customers</li>

<li>Passengers mostly take short distance travel (below 1000 km) and it contributes to the high number of dissatisfaction.</li>

</ol>

## Reflections <a name="2.6"></a>

After performing the univariate and bivariate analysis, we can conclude that the airline service is not satisfying enough for the customers. The majority of passengers are not satisfied with the airline service, especially those who are first-time flyers and economy class passengers. The airline should focus on improving the services for these two segments to increase customer satisfaction. The analysis also shows that the airline has a strong returning customer base, which is a positive sign. However, the airline should not take this for granted and should continue to improve its services to retain these customers and attract new ones.

# Data Modeling <a name="3"></a>

Now, since the dataset is ready to be modeled, we can assign it a more descriptive name. In this section, we will apply two modeling techniques to achieve the objectives of our analysis:

1.  Dimension Reduction
2.  Classification Method
    
    2.1 Classification using Decision Tree Model
    
    2.2 Classification using Logistic Regression Model

## Dimension Reduction <a name="3.1"></a>

Now, we will perform PCA (Principal Component Analysis) to reduce the dimensionality of the dataset. PCA is a technique used to reduce the number of variables in a dataset while retaining as much information as possible. It does this by transforming the original variables into a new set of variables called principal components, which are linear combinations of the original variables. The first few principal components capture most of the variance in the data, allowing us to simplify the dataset without losing too much information. Furthermore, PCA can be used to reduce the dimensionality of the dataset while ensuring that we dont lose too much information

**PCA with Numeric Variable Only**

First, we will perform PCA on the numeric variables only. We will use the standardized data to perform PCA. The standardized data is created by removing the ID column and the log-transformed columns, and then scaling the remaining numeric variables.

```{r}
# Use the standardized data
data_standardized <- data %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z,Departure.Delay.Duration, Arrival.Delay.Duration,
            Age.log, Age.log.z)) %>% 
  select(where(is.numeric)) %>%
  scale(center = TRUE, scale = TRUE)

# Perform PCA
results <- prcomp(data_standardized)
summary(results)
```

Then, we will plot the PCA results to visualize the variance explained by each principal component. The scree plot shows the proportion of variance explained by each principal component. The result shows that the cummulative value already achieved 90% using 12 Principal Components as below.

```{r}
# Plot the PCA results
results$rotation[ , c(1:2) ]
```

Next, To focus on variable with high loadings, we set the threshold of 0.2 and only select the first principal component.

```{r}
# Loadings for the first principal component
loadings <- data.frame(results$rotation)
threshold <- 0.20
indices <- abs(loadings$PC1) >= threshold
data.frame(variable = rownames(loadings)[indices], 
           loading = loadings$PC1[indices])
```

Furthermore, we will plot the loadings of the first principal component to visualize which variables contribute most to the first principal component.

```{r}
# Get loadings for the first principal component
loadings_df <- data.frame(
  variable = rownames(loadings)[indices],
  loading = loadings$PC1[indices]
)

# Plot the loadings
ggplot(loadings_df, aes(x = reorder(variable, -loading), y = loading, fill = loading > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Top Loadings for PC1",
       x = "Variable",
       y = "Loading Value") +
  scale_fill_manual(values = c("steelblue", "tomato")) +
  theme_minimal()
```

The loadings represent the correlation between the original variables and the principal components. A higher absolute value indicates a stronger relationship with that component. Which is the top 5 contributors to the first principal component are flight entertainment, comfortability, in-flight entertainment, in-flight service, and cleanliness.

Next, we are analyzing top contributor in PC1-PC12

```{r}
# Get loadings for all components
loadings <- as.data.frame(results$rotation)
loadings$Variable <- rownames(loadings)

# Top contributors for PC1–PC12
top_vars <- apply(abs(loadings[, 1:12]), 2, function(x) names(sort(x, decreasing = TRUE)[1:5]))
print(top_vars)
```

PC1 – In-Flight Amenities & Comfort Driven by Flight.Entertainment, Comfortability, In.flight.Entertainment, Flight.general.service, and Cleanliness. This component captures the overall in-flight experience focused on comfort and entertainment.

PC2 – Digital & Booking Experience Influenced by Ease.of.Online.Booking, Pre.flight.service, and In.flight.Wifi.Service. It represents passengers’ interaction with digital services before and during the flight.

PC3 – Operational Onboard Service Dominated by In.flight.Service, Baggage.Handling, and Seat.Comfort, this component reflects service reliability and seat-related comfort during the flight.

PC4 – Flight Timing & Boarding Characterized by Arrival.Delay, Departure.Delay, and Online.Boarding. It relates to punctuality and boarding process efficiency.

PC5 – Passenger Profile & Journey Shaped by Online.Boarding, Flight.Distance, Age, and Gate.Location, this component reflects demographic and journey-specific traits.

PC6 – Schedule & Gate Convenience Led by Departure.and.Arrival.Time.Convenience, Gate.Location, and Age, this component captures how convenient flight times and gate accessibility are for passengers.

PC7 – Pre-Flight & Cabin Space Defined by Check.in.Service, Leg.Room.Service, and Flight.Distance, combining service before boarding with cabin comfort.

PC8 – Age-Based Travel Patterns Mainly driven by Age, Flight.Distance, and Check.in.Service, this component shows how age may influence travel experience and expectations.

PC9 – Comfort in Long Trips Includes Leg.Room.Service, Flight.Distance, and Comfortability, highlighting comfort needs for longer or more frequent flights.

PC10 – In-Flight Consumables Composed of On.board.Service, Flight.general.service, and Food.and.Drink, focusing on passengers’ satisfaction with in-flight consumables.

PC11 – Delays & Terminal Logistics Influenced by Arrival.Delay, Departure.Delay, and Gate.Location, this component represents disruptions tied to flight and gate logistics.

PC12 – Check-In & Time Coordination Driven by Departure.and.Arrival.Time.Convenience and Check.in.Service, highlighting the efficiency of early-stage travel processes.

```{r}
# Plot the scree plot
screeplot(results, type = "lines")
```

From the plot above, we can see that the first 12 components explain over 90% of the variance in the data. This indicates that the data is well-structured and that there are strong correlations between the variables. The first few components capture most of the variance in the data, allowing us to simplify the dataset without losing too much information.

**PCA with One Hot Encoding**

From the previous analysis, we can see that the PCA is performed on the numeric variables only. However, we can also perform PCA on the categorical variables by using one-hot encoding. One-hot encoding is a technique used to convert categorical variables into a format that can be provided to machine learning algorithms to do a better job in prediction.

```{r}
# Use the standardized data
data_standardized <- data %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z,Departure.Delay.Duration, Arrival.Delay.Duration,
            Age.log, Age.log.z))

# One-hot encode categorical variables
dummies <- dummyVars(~ ., data = data_standardized)
data_encoded <- predict(dummies, newdata = data_standardized)

# Standardize the data
data_encoded_standardized <- scale(data_encoded, center = TRUE, scale = TRUE)

# Perform PCA
results <- prcomp(data_encoded_standardized)
summary(results)
```

When PCA was applied to numeric-only variables, the first 12 components explained over 90% of the variance, indicating strong inter-variable correlation and compact data structure. However, after including one-hot encoded categorical variables, the number of components required to achieve the same level of variance increased to 20. This suggests that categorical variables contribute additional but more dispersed variance, requiring more dimensions to capture the overall structure.

## Classification Method : Decision Tree Model <a name="3.2"></a>

Classification is an example of supervised learning where the model is trained to predict label variables based on the features of the dataset. In this case, we will use decision tree as the classification method,as the most popular and widely used classification method. Decision tree is a flowchart-like structure where each internal node represents a test on an attribute, each branch represents the outcome of the test, and each leaf node represents a class label. It works by splitting the data into subsets based on the value of the input features. The splitting process continues until a stopping criterion is met, such as reaching a maximum depth or minimum number of samples in a leaf node.

We use c50 as the package to implement decision tree. C5.0 is an improved version of C4.5, which is a popular algorithm for generating decision trees. It is known for its speed and accuracy, and it can handle large datasets with many features. C5.0 also includes features such as boosting, which can improve the performance of the model by combining multiple weak classifiers into a strong classifier.

### Model Preparation <a name="3.2.1"></a>

```{r}
# Create copy of dataset for the purpose of classification only, convert to factor
data_clsf <- data %>%
      mutate(across(-c(ID, Age, Flight.Distance, Departure.Delay, Arrival.Delay, 
                   Departure.Delay.Duration,
                   Arrival.Delay.Duration, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
                   Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
                   Flight.Distance.log.z, Age.log, Age.log.z), as.factor))

# Encode Target Variable
data_clsf$Satisfaction <- ifelse(data_clsf$Satisfaction=="Satisfied",1,0)
data_clsf$Satisfaction <- factor(data_clsf$Satisfaction, levels = c(0, 1))

# For experiment, create subset of untransformed data, data with log, and data with normalization
data_clsf_ori <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z, Age.log, Age.log.z)) 

data_clsf_log <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log.z, Departure.Delay.Duration.log.z, Flight.Distance.log.z,
            Departure.Delay.Duration, Arrival.Delay.Duration, Age.log.z, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age)) 

data_clsf_norm <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log,
            Departure.Delay.Duration.log, Flight.Distance.log,
            Age.log, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age)) 
```

From the Model preparation, we can see that the dataset is ready to be used for classification. The target variable is **Satisfaction** which is a binary variable with 1 for satisfied and 0 for not satisfied. The other variables are the predictors which are all categorical variables. The dataset is split into three datasets: untransformed, log-transformed, and normalized.

**Split Dataset Training Testing for Untransformed Data**

```{r}
set.seed(46748717)

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)
```

### Training a decision tree model on the data <a name="3.2.2"></a>

From the training data, we will train a decision tree model using the C5.0 algorithm.

::: scroll-box
```{r}
# Train the model
model <- C50::C5.0(Satisfaction ~., 
                   data = training)

# Print the model summary
summary(model)
```
:::

From training a decision tree model, we can see that the model is trained with the training data. The summary of the model shows the number of rules and the number of leaves in the decision tree. The model is trained with 10 rules and 10 leaves.

Next, we visualize the model to better understand how the decision tree makes predictions.

Plotting top 5 model :

```{r fig.width = 10, fig.height = 10}

# Visualization-only version, targeting ~5 levels
model_top5 <- C5.0(Satisfaction ~ ., data = training,
                   control = C5.0Control(minCases = 3500, CF = 0.25))
plot(model_top5)
```

From the graph above,nodes are the ovals (decisions) and rectangles (final outcomes), branches are the lines connecting them with condition labels, and levels represent the depth from top (root) to bottom (leaves), where predictions are made. We can see that the most important factor influencing customer satisfaction is In-flight Wifi Service, as it is the first and most frequent split in the tree, indicating strong predictive power. Customers who rated Wifi poorly (values 0 or 5) are more likely to be dissatisfied (Node 2), while those who gave moderate to high ratings proceed through other factors. Among these, Online Boarding and Type of Travel play key roles, especially for business travelers, whose satisfaction is further influenced by In-flight Wifi Service again. Additionally, Class (e.g., Economy vs. Business) and Gate Location appear in deeper splits, showing they contribute but are less critical. Overall, Wi-Fi satisfaction, especially for business travelers, is the top predictor of overall satisfaction.

*In-flight service* , *Online boarding*, *Type of travel*, *Class*, and *Gate Location*.

```{r}
print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Model with data testing:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")

```

As we can see at the metrics for the model, it shows high accuracy at 96%. The model is able to predict the satisfaction level of the customers with high accuracy. The precision and recall are also high at 95% and 98% respectively, indicating that the model is able to correctly identify the satisfied customers and dissatisfied customers. The F1-score is also high, indicating that the model is able to balance precision and recall well.

**Try with dataset that already log transformed**

::: scroll-box
```{r}
set.seed(46748717)

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_clsf_log, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the model
model <- C50::C5.0(Satisfaction ~., 
                   data = training)

# Print the model summary
summary(model)

# Print the confusion matrix
print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Model with data training:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")

```
:::

The result is almost identical since the algorithm determine the transformed data is not have significant importance. Therefore the improvement will focus on tuning the parameter of the model.

### Model Improvement <a name="3.3.3"></a>

In this section, we tried to explore whether there is a room to improve the model performance. Note that the model performance is considered good with accuracy of 96%, but we will try to improve it by tuning the parameters of the model. The model performance is measured by accuracy, precision, recall, and F1-score. The model performance is evaluated using the testing data.

**1. Boosting**

**Boosting** is one of methods to improve the performance of the model. The idea is to combine multiple weak classifiers to create a strong classifier. In this section, we will use boosting to improve the performance of the decision tree model.Key concept of boosting includes sequential learning where each new model is trained to correct the errors of the previous model (Schwenk,2000). The final model then is a weighted sum of all the models, where the weights are determined by the performance of each model. Finally, loss minimization is used to minimize the error of the final model.

```{r}
set.seed(46748717)

# Split the data into 70% training and 30% testing
proportion <- 0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the boosting model
model_boost <- C5.0(Satisfaction ~ ., 
                    data = training, 
                    trials = 10)

# Predict on the testing data
pred.test <- predict(model_boost, testing)

# Generate the confusion matrix
print("Confusion matrix based on testing data (boosting)")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Boosting Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")

```

Let's compare with the previous model performance

| Metric    | Previous Model | Boosting Model |
|-----------|----------------|----------------|
| Accuracy  | 0.96           | 0.96           |
| Precision | 0.95           | 0.96           |
| Recall    | 0.98           | 0.97           |
| F1-Score  | 0.97           | 0.96           |

Based on the result above, there is no significant improvement in terms of model performance. This strengthen the idea that the model performance is already good. However, we can see that the precision is slightly improved from 0.95 to 0.96, while the recall is slightly decreased from 0.98 to 0.97. The F1-score is also slightly decreased from 0.97 to 0.96. In conclusion, our model performance is already good and there is no significant improvement after boosting.

**2. Assigning Cost to Mistake**

Now let's try another method to improve the model which is **Assigning cost to mistake**. this method is used to assign different costs to different types of mistakes. This involves setting optimal costs for different classes to suppress the dominance of majority classes during training, thereby enhancing the model's diagnostic capabilities (Ren, 2022). In this case, we will assign a higher cost to predicting a satisfied customer as dissatisfied and a lower cost to predicting a dissatisfied customer as satisfied. This is because the cost of losing a satisfied customer is higher than the cost of losing a dissatisfied customer. The cost matrix is defined as follows:

```{r}
# Specifying the cost matrix
cost.matrix <- matrix(c(NA, 1,  # FN costs of predicting "Disatisfied" whereas actual value is "Satisfied"
                        2, NA), # FP costs of predicting "Satisfied" whereas actual value is "Disatisfied"
                      nrow = 2, 
                      ncol = 2,
                      byrow = FALSE)
rownames(cost.matrix) <- colnames(cost.matrix) <- c(1, 0)

# Estimating the model with the cost matrix
model.cost <- C5.0(Satisfaction ~., 
              data = training,
              costs = cost.matrix)

# Predicting on the testing data
pred.test <- predict(model.cost, testing)

# Generating the confusion matrix
print("Confusion matrix based on testing data (costs")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Cost sensitive learning  Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

Let's compare with the previous model performance

| Metric    | Previous Model | Cost Sensitive Model |
|-----------|----------------|----------------------|
| Accuracy  | 0.96           | 0.96                 |
| Precision | 0.95           | 0.94                 |
| Recall    | 0.98           | 0.99                 |
| F1-Score  | 0.97           | 0.96                 |

Based on the performance comparison, cost mistake assignment result is not significantly different with the model. This indicates that our model performance is having a good performance already. However, we can see that the recall is slightly increased from 0.98 to 0.99, while the precision is slightly decreased from 0.95 to 0.94. The F1-score is also slightly decreased from 0.97 to 0.96. In conclusion, our model performance is already good and there is no significant improvement after assigning cost to mistake.

**3. Random Forest**

Another method for model improvement is **Random Forest**. This method is an ensemble learning method that combines multiple decision trees to create a more robust model. The idea is to create multiple decision trees using different subsets of the data and then combine the predictions of each tree to make the final prediction. This method is less prone to overfitting and can handle high-dimensional data better than a single decision tree (Biau, 2010). Key features of random forest include:<br>

<ol>

<li>Diversity: each tree is trained on a random subset of data and features, ensuring diversity</li>

<li>Robustness: It reduces the risk of overfitting (common in single decision trees) by averaging the results from many trees.</li>

<li>Feature Importance: Random Forest ranks the importance of features, helping you understand which variables contribute most to predictions.</li>

</ol>

```{r random forest}
# Generate dataset for random forest
data_randfor <- data_clsf_ori %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age)) 

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix based on testing data (costs")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Random Forest  Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

Let's compare with the previous model performance

| Metric    | Previous Model | Random Forest Model |
|-----------|----------------|---------------------|
| Accuracy  | 0.96           | 0.95                |
| Precision | 0.95           | 0.95                |
| Recall    | 0.98           | 0.97                |
| F1-Score  | 0.97           | 0.96                |

Based on the performance comparison, random forest result provides not significant differences with the current model. This is because our model performance is already showing a good performance thus the room to improve is small. However, we can see that the accuracy is slightly decreased from 0.96 to 0.95, while the precision is unchanged at 0.95. The recall is slightly decreased from 0.98 to 0.97. The F1-score is also slightly decreased from 0.97 to 0.96. In conclusion, our model performance is already good and there is no significant differences after using random forest.

**Using Ranger to implement random forest to the database with high-cardinality factors**

Last method to improve is using random forest with the consideration of **high cardinality** data using ranger. The data used in this case can be considered having high cardinality since the data is categorical and has many levels. The ranger package is a fast implementation of random forests that can handle high-dimensional data efficiently (Wright, 2015). It is particularly useful for large datasets with many features, as it can quickly build and evaluate random forest models.<br>

Why high cardinality matters? <br>

High cardinality categorical variables can lead to **overfitting** in traditional decision tree algorithms.This is because the model may create many splits based on rare categories, which can reduce the model's generalization ability. Ranger uses a different approach to handle high cardinality data by using a technique called "feature selection" to reduce the number of features used in the model. This helps to prevent overfitting and improve the model's performance.

```{r}
# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the ranger model
model.ranger <- ranger(Satisfaction ~., 
                       data = training,
                       num.trees = 500, # how many trees should be grown?
                       mtry = 2, # how many variables to sample at each split?
                       replace = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.ranger, testing)$predictions

# Print the confusion matrix
print("Confusion matrix based on testing data (costs")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics for Random Forest with Ranger  Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

Let's compare with the previous model performance

| Metric    | Previous Model | Random Forest with Ranger Model |
|-----------|----------------|---------------------------------|
| Accuracy  | 0.96           | 0.95                            |
| Precision | 0.95           | 0.95                            |
| Recall    | 0.98           | 0.97                            |
| F1-Score  | 0.97           | 0.96                            |

The result of this method is similar with the previous random forest method. There is no significant changes in the model performance. This is mainly because the model is already showing a good performance leaving a small room to improve. However, accuracy slightly decreased from 0.96 to 0.95, while the precision is unchanged at 0.95. The recall is slightly decreased from 0.98 to 0.97. The F1-score is also slightly decreased from 0.97 to 0.96. In conclusion, our model performance is already good and there is no significant differences after using random forest with ranger.<br>

**To conclude, we can say that the model performance is already good and there is no significant improvement after using boosting, assigning cost to mistake, random forest, and ranger.**

### Model Evaluation <a name="3.4"></a>

We have developed the model using classification method. In this section, we will try to evaluate the model using variable importance.<br>

Evaluation of variable importance is crucial in business analytics as it helps to identify the most significant factors influencing customer satisfaction. By understanding which variables have the greatest impact, businesses can prioritize their efforts and resources to improve those areas, ultimately enhancing customer satisfaction. In this case we will analyze the variable importance using the random forest model. Random forests provide a built-in mechanism to assess the importance of each feature in making predictions. This is done by measuring how much each feature contributes to reducing the overall prediction error.

```{r}
# Print the variable importance
importance <- importance(model.forest)
print(importance)
```

Above is the importance of each variable in the model. The higher the value, the more important the variable is in predicting customer satisfaction. There are two main parameters observed in the importance evaluation, they are:<br>

<ol>

<li>

**Mean Decrease Accuracy (MDA)**: This measures the decrease in model accuracy when the variable is permuted. A higher value indicates that the variable is important for making accurate predictions.<br> How to interpret MDA: <br>

<ul>

<li>High MDA: The variable is crucial for the model's predictive power. If you remove or permute this variable, the model's accuracy drops significantly.</li>

<li>Low MDA: The variable has little impact on the model's predictions. Removing it won't significantly affect the model's performance.</li>

</ul>

</li>

<li>

**Mean Decrease Gini (MDG)**: This measures the total decrease in node impurity (Gini impurity) brought by the variable across all trees in the forest. A higher value indicates that the variable is important for splitting nodes and reducing uncertainty.<br> How to interpret MDG: <br>

<ul>

<li>High MDG: The variable is effective in splitting the data into homogeneous groups. It plays a significant role in the decision-making process of the model.</li>

<li>Low MDG: The variable is less effective in splitting the data. It may not provide much information for the model's predictions.</li>

</ul>

</li>

</ol>

Let's visualize the variable importance using a bar plot. The plot will show the MDA and MDG values for each variable, allowing us to easily identify the most important features in the model.

```{r varimp-plot, fig.width=10, fig.height=8}

# Print the Variable Importance Plot
varImpPlot(
  model.forest,
  main = "Variable Importance for Random Forest Model",
   pch = 16 # Use filled circles for data points
)

```

Based on the graph, we can see that the most important features are: <br> **MDA:**<br>

<ol>

<li>Type of travel</li>

<li>Inflight wifi service</li>

<li>Customers type</li>

</ol>

these features have the highest importance. This suggests that these features significantly impact the classification performance.<br> Features with low or near-zero values (e.g., arrival delay status, departure delay status, delay status) contribute less and might be candidates for exclusion during model optimization.<br>

**MDG:**<br>

<ol>

<li>Online boarding</li>

<li>Inflight wifi service</li>

<li>type of travel</li>

</ol>

These features show the greatest contributions to reducing impurity.<br> The features with low or near-zero values (e.g., departure delay status, arrival delay status, delay status) contribute less and considered not significant.<br> <br>

**Importance Evaluation on high satisfaction VS Low satisfaction** <br> Next, we will try to take a look on importance in high satisfaction and low satisfaction. We will use the same method as above to evaluate the importance of each variable in high satisfaction and low satisfaction. This phase is to evaluate the importance profile on both high satisfaction and low satisfaction. The goal is to identify the most important features that contribute to high satisfaction and low satisfaction. This will help us to understand the factors that influence customer satisfaction and dissatisfaction.<br><br>

**Step 1: high satisfaction variable importance evaluation**<br> High satisfaction category based on data:<br>

<ol>

<li>Type of travel : Business</li>

<li>Class : Business</li>

<li>Customer Type : Returning</li>

</ol>

**Notice** that in the fist plot we see non-satisfaction contributors being plotted. In the following step we will focus on possible contributors to the satisfaction by evaluating variable importance.<br> <br>

<span style="color: #1E90FF; font-size: 16px;">**Type of Travel : Business**</span><br>

```{r fig.width=10, fig.height=8}
# Subset the data on passenger with Business type of travel
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Business"), ]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data 
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Type of Travel Business")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")

```

```{r fig.width=10, fig.height=8}

# Plot the variable importance
varImpPlot(
  model.forest,
  main = "Variable Importance for Type of Travel: Business",
   pch = 16 # Use filled circles for data points
)

```

<span style="color: #1E90FF; font-size: 16px;">**Class: Business**</span><br>

```{r}
# Subset the data on passenger with Business Class flight
data_randfor <- data_clsf_ori[which(data_clsf_ori$Class == "Business"),]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Class Business")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

```{r fig.width=10, fig.height=8}

# Plot the variable importance
model.forest.business <- model.forest

varImpPlot(
  model.forest.business,
  main = "Variable Importance for Class: Business",
   pch = 16 # Use filled circles for data points
)

```

<span style="color: #1E90FF; font-size: 16px;">**Customers Type : Returning**</span><br>

```{r fig.width=10, fig.height=8}
# Subset the data on passenger with returning customer type
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type == "Returning"), ]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Customer Type Returning")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

```{r fig.width=10, fig.height=8}
varImpPlot(
  model.forest,
  main = "Variable Importance for Customers Type: Returning",
   pch = 16 # Use filled circles for data points
  )
```

**Step 2: Low satisfaction variable importance evaluation**<br> Low satisfaction category based on data:<br>

<ol>

<li>Type of travel : Personal</li>

<li>Class : Economy and Economy Plus</li>

<li>Customer Type : First Time</li>

</ol>

<span style="color: #1E90FF; font-size: 16px;">**Type of Travel: Personal**</span><br>


```{r}
# Subset the data on passenger with Business Class flight
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Personal"), ]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Type of Travel Personal")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

```{r fig.width=10, fig.height=8}
# Plot the variable importance
varImpPlot(
  model.forest,
  main = "Variable Importance for Type of Travel: Personal",
   pch = 16 # Use filled circles for data points
  )
```

<span style="color: #1E90FF; font-size: 16px;">**Class: Economy and Economy Plus**</span><br>

```{r}
# Subset the data on passenger with Economy and Economy Plus Class flight
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Class %in% c("Economy", "Economy Plus")), ]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Type of Class Economy and Economy Plus")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

```{r fig.width=10, fig.height=8}
# Plot the variable importance

model.forest.economy <- model.forest

varImpPlot(
  model.forest.economy,
  main = "Variable Importance for Class: Economy and Economy Plus",
   pch = 16 # Use filled circles for data points
  )
```

<span style="color: #1E90FF; font-size: 16px;">**Customers Type : First time**</span><br>

```{r}
# Subset the data on passenger with Business Class flight
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type %in% "First-time"), ]
data_randfor <- data_randfor %>% select(-c(Type.of.Travel, Gender, Customer.Type, Age_Range, Distance_Group, Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age, Class))

# Split the data into 70% training and 30% testing
proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

# Train the random forest model
model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

# Predicting on the testing data
pred.test <- predict(model.forest, testing)

# Print the confusion matrix
print("Confusion matrix for Random Forest Type of Customer Type First Time")
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
# --- Calculate Metrics ---

# Create the confusion matrix table
conf_matrix <- table(Actual = testing$Satisfaction, Predicted = pred.test)

# Extract True Positives, True Negatives, False Positives, and False Negatives. 0 representing dissatisfaction, 1 represent satisfaction.
tp <- conf_matrix["0", "0"]
tn <- conf_matrix["1", "1"]
fp <- conf_matrix["1", "0"]
fn <- conf_matrix["0", "1"]

# Calculate metrics
accuracy <- (tp + tn) / sum(conf_matrix)
precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

# Print metrics
cat("\nMetrics of Model:\n")
cat("Accuracy: ", round(accuracy, 2), "\n")
cat("Precision: ", round(precision, 2), "\n")
cat("Recall: ", round(recall, 2), "\n")
cat("F1-Score: ", round(f1_score, 2), "\n")
```

```{r fig.width=10, fig.height=8}
# Plot the variable importance
varImpPlot(
  model.forest,
  main = "Variable Importance for Customers Type: First time",
   pch = 16 # Use filled circles for data points
  )
```

Let's evaluate the variable importance result. The comparison between Top MDA on high satisfaction and low satisfaction customers will help to understand what parameters matter to define their satisfaction level. Below is the comparison between high satisfaction and low satisfaction customers based on the top MDA value. The table below shows the top 3 parameters that matter to high satisfaction and low satisfaction customers. The table is divided into 3 categories, which are: Type of Travel, Class, and Customers Type.<br> <br>

| **Parameters Top MDA** | **High Satisfaction** | **Low Satisfaction** |
|-----------------------|-----------------------|---------------------------|
| **Type of Travel** | **Business** | **Personal** |
| Top 1 | Online Boarding | In-flight WiFi Service |
| Top 2 | Check-in Service | Ease of Online Booking |
| Top 3 | Baggage Handling | Flight Entertainment |
| **Class** | **Business** | **Economy and Economy Plus** |
| Top 1 | Online Boarding | In-flight WiFi Service |
| Top 2 | Departure and Arrival Time Convenience | Baggage Handling |
| Top 3 | Check in Service | Ease of online booking |
| **Customers Type** | **Returning** | **First time** |
| Top 1 | Departure and Arrival Time Convenience | In flight wifi service |
| Top 2 | Online Boarding | Baggage Handling |
| Top 3 | In flight wifi service | In flight service |

<br> The table above shows that the top 3 parameters that matter to high satisfaction customers are different from the low satisfaction customers. <br> **Overall Insight** For high satisfaction, the recurring themes are efficiency, convenience, and timeliness. Business travel and returning customers emphasize this the most. For low satisfaction, comfort, accessibility, and entertainment take precedence, especially for personal travel and first-time customers.<br>

**Let's take a look on each group** For **Type of Travel**, high satisfaction is driven by business travelers who value efficiency and convenience, particularly through services like online boarding, check-in service, and baggage handling. In contrast, low satisfaction among personal travelers stems from a focus on comfort and accessibility, with preferences for in-flight WiFi service, ease of online booking, and flight entertainment.<br>

For **Class**, high satisfaction is prominent among Business Class customers who emphasize seamless travel experiences, prioritizing online boarding, departure and arrival time convenience, and check-in service. On the other hand, low satisfaction in Economy and Economy Plus passengers highlights their need for enjoyable and accessible features, such as in-flight WiFi service, baggage handling, and ease of online booking.<br>

For **Customer Type**, returning customers experience high satisfaction when services like departure and arrival time convenience, online boarding, and in-flight WiFi service cater to their need for smooth and timely travel. Meanwhile, low satisfaction among first-time customers arises from their focus on pleasant first-time experiences, valuing in-flight WiFi service, baggage handling, and in-flight service.<br>

However, we can see the convergence of the results between high satisfaction and low satisfaction customers.<br>
For business-class passengers, the key contributors to high satisfaction—such as Online Boarding, Check-in Service, and Departure and Arrival Time Convenience—show consistency across categories (e.g., travel type and class). This indicates that factors driving satisfaction are aligned, supporting the idea that a single model for the **Business segment** is sufficient. <br>
For economy-class passengers (including Economy Plus), dissatisfaction revolves around similar factors like In-flight WiFi Service, Ease of Online Booking, and Baggage Handling across different categories. This suggests shared pain points that can be effectively captured in a unified model for the **Economy segment.**<br>
This convergence implies that a single model for each segment (Business and Economy) can effectively address the unique needs of passengers, enhancing satisfaction and reducing dissatisfaction. <br><br>
Reducing the models from six to two avoids over-complication and overfitting while ensuring that the key insights remain actionable. As per data, it is shown that the segmentation between Business and Economy is quite competitive, around 52% of economy combined (including economy plus) and 48% of business segment. **Next we will perform logistic regression with the two main classification of Business segment and Economy segment**.<br><br>
Logistic regression works well when models are representative and not overly complex. Consolidating into two models ensures each has enough data points and avoids redundancy, improving the reliability of your conclusions (Landwehr, 2003).


## Logistic Regression <a name="3.3"></a>

Since this project aims to predict between binary class values (satisfied and dissatisfied), logistic regression will be used for this classification task (Hosmer, 2013). This method also allows us to identify the impact of each variable to the likelihood of passenger satisfaction, as well as see the probability values associated. In this subchapter, we will discuss step-by-step our process of preparing the regression model as well as interpret our findings.

### Model Preparation <a name="3.3.1"></a>

Firstly, we will prepare the dataset by applying a train-test split of 70:30. We will use seed value to get replicable results.

```{r}
set.seed(46748756)

# Treat ordinal variables as numeric and exclude specified variables
data_numeric <- data %>%
  mutate(across(where(is.numeric), ~ scale(.) %>% as.numeric())) %>% # Normalize all numeric columns
  mutate(across(c(Ease.of.Online.Booking, Check.in.Service, Online.Boarding, 
                  Gate.Location, On.board.Service, Seat.Comfort, Leg.Room.Service, 
                  Cleanliness, Food.and.Drink, In.flight.Service, In.flight.Wifi.Service, 
                  In.flight.Entertainment, Departure.and.Arrival.Time.Convenience), 
                as.numeric)) %>% # Treat ordinal variables as numeric
  subset(select = -c(ID, Gender, Age, Customer.Type, Type.of.Travel, Flight.Distance, 
                     Age_Range, Distance_Group, Age.log, Flight.Distance.log.z, Age.log.z, 
                     Departure.Delay.Duration, Arrival.Delay.Duration, 
                     Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z, 
                     Departure.Delay.Duration.log, Departure.Delay.Duration.log.z,
                     Flight.Distance.log, Flight.general.service, Flight.Entertainment,
                     Pre.flight.service, Comfortability)) # Exclude unnecessary columns

# Cleaning columns with zero scores that exist in small quantities
cols_to_clean <- c("On.board.Service", "Check.in.Service", "Seat.Comfort",
                   "In.flight.Service", "Gate.Location", "Cleanliness", 
                   "In.flight.Entertainment", "Flight.general.service",
                   "Flight.Entertainment", "Pre.flight.service", "Comfortability")

for (col in colnames(data_numeric)) {
  if (is.factor(data_numeric[[col]])) {
    data_numeric[[col]] <- droplevels(data_numeric[[col]])
  }
}

# Split the dataset into training and testing sets
proportion <- 0.7
split <- initial_split(data_numeric, prop = proportion)
training <- training(split)
testing <- testing(split)

# Encode Satisfaction as binary (0 = Neutral or Dissatisfied, 1 = Satisfied)
training$Satisfaction <- ifelse(training$Satisfaction == "Satisfied", 1, 0)
training$Satisfaction <- factor(training$Satisfaction, levels = c(0, 1))
testing$Satisfaction <- ifelse(testing$Satisfaction == "Satisfied", 1, 0)
testing$Satisfaction <- factor(testing$Satisfaction, levels = c(0, 1))

# Split the dataset by class
business_class_data <- training %>% filter(Class == "Business") %>% select(-Class)
economy_class_data <- training %>% filter(Class %in% c("Economy", "Economy Plus")) %>% select(-Class)

```

The code chunk above first converts all specified columns into factors, as logistic regression requires categorical predictors to be factors. Next we remove one row which contains zero column values, as this cannot be split into the training and testing datasets. Finally, the dataset is split and we extract the training and testing set.

### Training Logistic Regression Model <a name="3.3.2"></a>

Once we have determined the training and testing set, we will train the model in the code chunk below.

-   **`glm()`**: Fits a generalized linear model.

-   **`Satisfaction ~ .`**: Specifies the formula, where `Satisfaction` is the dependent variable, and all other variables in the `training` dataset are independent variables.

-   **`data = training`**: Uses the `training` dataset to train the model.

-   **`family = "binomial"`**: Specifies that the model is a logistic regression, as the dependent variable (`Satisfaction`) is binary.

Three models will be trained, the first one including all the data which we will use to compare with the classification model and as a baseline, the second and third models are used to identify differing characteristics between passengers of business class and economy class.

```{r training_logreg_alldata}

# Using all data
model_log <- glm(Satisfaction ~., 
               data = training, family = "binomial")
summary(model_log)
```

```{r training_logreg_business}

# Using business class data
model_log_business <- glm(Satisfaction ~., 
               data = business_class_data, family = "binomial")
summary(model_log_business)
```

```{r training_logreg_economy}

# Using economy class data
model_log_economy <- glm(Satisfaction ~., 
               data = economy_class_data, family = "binomial")
summary(model_log_economy)
```

The results above show the relationship between the predictor variables and the target variables. We are interested in determining the significant predictors of the model. In the next subchapter, we will delve deeper into the interpretation of these results.

### Interpretation of the Model <a name="3.3.3"></a>

**Interpret the coefficient**

The code chunk below is used to calculate the exponentiated coefficients (also known as odds ratios) from the logistic regression model `model_log`.

```{r}
exp(coef(model_log))
```

```{r}
exp(coef(model_log_business))
```

```{r}
exp(coef(model_log_economy))
```

The odds ratio indicates how the odds of the target variable (e.g., `Satisfaction` being "Satisfied") change with a one-unit increase in the predictor variable, holding all other variables constant.

-   **Odds Ratio \> 1**: The predictor increases the odds of satisfaction.

-   **Odds Ratio \< 1**: The predictor decreases the odds of satisfaction.

-   **Odds Ratio = 1**: The predictor has no effect on satisfaction.

We find that strong positive predictors include returning customers, business travelers, ease of online booking, and older passengers. Meanwhile, negative predictors include economy class travelers, low rating of online boarding and delays.

**Visualizing the coefficients**

Below we will visualize the magnitude of each predictor's effect on the target variable, along with their 95% confidence intervals. The X axis represents the coefficient estimates in log-odds, horizontal lines around each point represent the confidence intervals (at 95% level).

```{r fig.width=10, fig.height=20}

coefs <- summary(model_log)$coefficients

coef_df <- data.frame(
  Variable = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  StdError = coefs[, "Std. Error"],
  p_value = coefs[, "Pr(>|z|)"]
)

# Remove intercept for clarity
coef_df <- coef_df[coef_df$Variable != "(Intercept)", ]

# Plot using ggplot2
ggplot(coef_df, aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError,
                    ymax = Estimate + 1.96 * StdError), width = 0.2) +
  coord_flip() +
  labs(title = "Logistic Regression Coefficients (All Class)",
       x = "Predictor",
       y = "Coefficient Estimate (log-odds)") +
  theme_minimal()
```

```{r fig.width=10, fig.height=20}

coefs <- summary(model_log_business)$coefficients

coef_df <- data.frame(
  Variable = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  StdError = coefs[, "Std. Error"],
  p_value = coefs[, "Pr(>|z|)"]
)

# Remove intercept for clarity
coef_df <- coef_df[coef_df$Variable != "(Intercept)", ]

# Plot using ggplot2

ggplot(coef_df, aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError,
                    ymax = Estimate + 1.96 * StdError), width = 0.2) +
  coord_flip() +
  labs(title = "Logistic Regression Coefficients (Business Class)",
       x = "Predictor",
       y = "Coefficient Estimate (log-odds)") +
  theme_minimal()
```

```{r fig.width=10, fig.height=20}
coefs <- summary(model_log_economy)$coefficients
coef_df <- data.frame(
  Variable = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  StdError = coefs[, "Std. Error"],
  p_value = coefs[, "Pr(>|z|)"]
)
# Remove intercept for clarity
coef_df <- coef_df[coef_df$Variable != "(Intercept)", ]

# Plot using ggplot2
ggplot(coef_df, aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError,
                    ymax = Estimate + 1.96 * StdError), width = 0.2) +
  coord_flip() +
  labs(title = "Logistic Regression Coefficients (Economy Class)",
       x = "Predictor",
       y = "Coefficient Estimate (log-odds)") +
  theme_minimal()
```

In the general model, we can see that economy class has the largest negative coefficient. While this may be good for enhancing overall accuracy, it may lead to non-operationalizable results in the business context as Economy and Business class passengers have different pain points. In the models for economy and business class, we see that passengers for each class have different value preferences. Next, we will take a look at the Z-values to see the significant predictors.

**Interpreting Z Value**

```{r}
model_summary_business <- summary(model_log_business)
model_summary_economy <- summary(model_log_economy)

# Extract z-values and convert to data frame
z_df_business <- data.frame(
  Variable_business = rownames(model_summary_business$coefficients),
  Z_value_business = model_summary_business$coefficients[, "z value"]
)

# Remove intercept
z_df_economy <- data.frame(
  Variable_economy = rownames(model_summary_economy$coefficients),
  Z_value_economy = model_summary_economy$coefficients[, "z value"]
)

# Remove intercept
z_df_business <- subset(z_df_business, Variable_business != "(Intercept)")
z_df_economy <- subset(z_df_economy, Variable_economy != "(Intercept)")
```

```{r  fig.width=10, fig.height=20}
# Combine the two dataframes into one for plotting
z_df_business$Group <- "Business Class"
z_df_economy$Group <- "Economy + Economy Plus"

# Rename columns for consistency
z_df_business <- z_df_business %>% rename(Variable = Variable_business, Z_value = Z_value_business)
z_df_economy <- z_df_economy %>% rename(Variable = Variable_economy, Z_value = Z_value_economy)

# Combine the dataframes
z_df_combined <- rbind(z_df_business, z_df_economy)


# Plot using ggplot2

ggplot(z_df_combined, aes(x = reorder(Variable, abs(Z_value)), y = Z_value, fill = Group)) +
  geom_col(position = "dodge") + # Side-by-side bars for each group
  geom_hline(yintercept = c(-1.96, 1.96), linetype = "dashed", color = "red") + # 95% significance threshold
  coord_flip() +
  labs(title = "Z-values from Logistic Regression by Class",
       x = "Predictor Variable",
       y = "Z-value (Wald Statistic)",
       fill = "Passenger Class") + # Add legend for the groups
  theme_minimal() +
  scale_fill_manual(values = c("Business Class" = "steelblue", "Economy + Economy Plus" = "green"))
```

The Z**-**values represent the Wald statistics for each predictor in the logistic regression model. These values are used to test the null hypothesis that the coefficient for a predictor is equal to zero (Agresti, 2018).

-   Positive Z-values indicate a **positive** relationship to the likelihood of satisfaction.

-   Negative Z-values indicate a **negative** relationship to the likelihood of satisfaction.

For a 95% confidence level, the threshold of a statistically significant Z-value is **\|Z\| \> 1.96** which is indicated by the dotted red line in the plot. We can see from the visualization above that the strongest negative predictor for business class is departure time and arrival convenience, while for economy class is ease of online booking. The strongest positive predictor for business class is online boarding, and for economy class it is in-flight WiFi. Next, we will evaluate the model.

### Model Evaluation <a name="3.3.4"></a>

The first calculation method that we will use to evaluate the logistic regression model is the confusion matrix. It is is a table that describes the performance of a classification model by comparing the predicted values with the actual values. It provides insight into how well the model is performing and helps identify any misclassifications. The result will display how many variables are correctly predicted as true positive and true negative as well as wrongly classified as false positive and false negative. The confusion matrix also calculate the accuracy, sensitivity, and specificity of the model. The accuracy is the proportion of true results (both true positives and true negatives) among the total number of cases examined. Sensitivity is the proportion of true positives that are correctly identified, while specificity is the proportion of true negatives that are correctly classified.

We will use the confusion matrix to evaluate the logistic regression model that includes all the data (business and economy class passengers), dubbed All Class. The logistic regression model for each economy class and business class is only used as a separator to provide the significant variables that affect the satisfaction level of each class. 

```{r}
all_testing <- testing
pred.test <- predict(model_log, newdata = all_testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- ifelse(testing$Satisfaction == 1, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- factor(all_testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

# Confusion matrix
confusionMatrix(pred.test.result, all_testing$Satisfaction)
```

The confusion matrix for combined class showed a relatively good result in all three aspects, with accuracy, sensitivity, and specificity are all well above 80%. To reconfirm the accuracy of the model, we will also check the ROC and AOC of the model.

**Interpret ROC and AUC**

Another quantified way to evaluate the model is to use the ROC curve and Area Under the Curve (AUC) value. The ROC curve is a graphical representation of the true positive rate (sensitivity) against the false positive rate (1 - specificity) at various threshold settings. AUC provides a single measure of overall model performance, with values ranging from 0 to 1. If the AUC value is closer to 1, the more accurate the model's prediction is. AUC value of 0.5 indicates that the model has no identified pattern of guessing, which is essentially the same as random guessing. Meanwhile, AUC value of 0 means that the model is completely wrong in its predictions.

We will now look at the ROC curve for the combined model (economy and business class) and its AUC value.

```{r}
# Running the model for all data
all_testing <- testing
pred.test <- predict(model_log, newdata = testing, type = "response")

# ROC curve and AUC
roc_obj <- roc(testing$Satisfaction, pred.test)
plot(roc_obj, col = "blue", main = "ROC Curve")
auc(roc_obj)
```

The ROC curve of All Class aboves shows that the model quickly reaches the value close to 1 in its initial phase, which indicates that the model has a good accuracy number. This is confirmed by the AUC value of 0.91, a value that is close to 1 and considered a good model.

### Improving Model Performance

**Using significant variables only**

Before we try several methods to improve the performance, we will display the summary of All Class model to see the AIC value. AIC (Akaike Information Criterion) is a measure of the relative quality of a statistical model for a given dataset. It is used to compare different models and select the best one. The lower the AIC value, the better the model fits the data. In this case, we will use AIC to compare the performance of different models.

```{r}
summary(model_log)
```

Above we see that the AIC value is 70515. We will use this value as a baseline to compare base model performance to the improved model.

The first method that we use to try improving the model performance is by filtering out the non-significant variables. We will use the p-value to determine the significance of each variable. The p-value is a statistical measure that helps to determine the significance of results. In this case, we will use a threshold of 0.05, which means that we will only keep the variables with a p-value less than 0.05. This is a common threshold used in statistical analysis to determine whether a result is statistically significant.

Below we will exclude the variables with p-value greater than 0.05, then we will also check using the `summary` function to see the AIC value.

p-value \< 0.05

```{r}
training_signif <- training %>% subset(select = -c(Departure.Delay, Arrival.Delay,
                                                  In.flight.Wifi.Service, Food.and.Drink, 
                                                  In.flight.Entertainment))

model_log_signif <- glm(Satisfaction ~., 
               data = training_signif, family = "binomial")

summary(model_log_signif)

```

We saw a slight increase in AIC value to 73629. Because the value is higher, it indicates that the model with filtered variables is actually not better than the base model for All Class model.

Next, we will evaluate the new model through the Confusion Matrix:

```{r}
pred.test <- predict(model_log_signif, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- ifelse(testing$Satisfaction == 1, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- factor(all_testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, all_testing$Satisfaction)
```

From the result above, the accuracy and sensitivity of the model with filtered variables experience a noticeable drop in accuracy and sensitivity value, while the specificity value is essentially the same. Again, this shows that the new model is not better than the base model.

Another method to evaluate the improvement attempt we want to use is using the ANOVA test to compare the two models. The ANOVA test is a statistical method used to compare the means of two or more groups. In this case, we will use the ANOVA test to compare the performance of the base model and the model with filtered variables. The ANOVA test will help us determine whether there is a significant difference between the two models. This is to ensure that the base model is indeed better than the model with filtered variables as indicated from the summary function and confusion matrix.

ANOVA test :

```{r}
anova(model_log, model_log_signif, test="Chisq")
```

The ANOVA test result above shows that the p-value is very close to 0, which indicates that there is a significant difference between the two models. This means that the base model is indeed better than the model with filtered variables, confirming the result from summary function and the confusion matrix.

**Using High Information Value Variables**

The second way to try improving the model is using the variables with High Information Value. Using the `Information` library, we can calculate the Information Value (IV) of each variable in the dataset. The IV is a measure of how much information a variable provides about the target variable. In other words, Information Value is a measure of the predictive capability of variables. Higher IV indicates that the variable is more predictive of the target variable. We will use this information to filter the variables and create a new model.

```{r}
library(Information)
library(dplyr)

# Step 1: Drop columns that contain "Date" or have more than one class
filtered_data <- data %>%
  select(-Satisfaction) %>%
  select(where(~ all(!grepl("Date|POSIX", class(.))) & length(class(.)) == 1))

# Step 2: Convert to numeric where possible
clean_data <- filtered_data %>%
  mutate(across(everything(), ~ if (all(!is.na(as.numeric(as.character(.x))))) {
    as.numeric(as.character(.x))
  } else {
    .x
  }))

# Step 3: Add binary target
clean_data$Satisfaction <- ifelse(data$Satisfaction == "Satisfied", 1, 0)

# Step 4: Run IV analysis
iv_results <- create_infotables(data = clean_data, y = "Satisfaction", bins = 10)

# View summary
iv_results$Summary

```

We will use the IV evaluation results to filter out the variables with insignificant value, with the threshold of the IV is 0.3. Everything below that value is considered weak predictor and will be excluded in the improvement model. We will use this information to filter the variables and create a new model, then evaluates the new model using confusion matrix and ANOVA test.

```{r}
iv_data <- data %>%
  mutate(across(where(is.numeric), ~ scale(.) %>% as.numeric())) %>% # Normalize all numeric columns
  mutate(across(c(Ease.of.Online.Booking, Check.in.Service, Online.Boarding, 
                  Gate.Location, On.board.Service, Seat.Comfort, Leg.Room.Service, 
                  Cleanliness, Food.and.Drink, In.flight.Service, In.flight.Wifi.Service, 
                  In.flight.Entertainment, Departure.and.Arrival.Time.Convenience), 
                as.numeric)) %>% # Treat ordinal variables as numeric
  subset(select = -c(ID, Gender, Gate.Location, Arrival.Delay, Departure.and.Arrival.Time.Convenience,
                     delay.status, arrival.delay.status, Departure.Delay, departure.delay.status,
                     Age_Range, Distance_Group, Age.log, Flight.Distance.log.z, Age.log.z, 
                     Departure.Delay.Duration, Arrival.Delay.Duration, 
                     Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z, 
                     Departure.Delay.Duration.log, Departure.Delay.Duration.log.z,
                     Comfortability, Flight.general.service, Flight.Entertainment,
                     Pre.flight.service)) # Exclude unnecessary columns

# Cleaning columns with zero scores that exist in small quantities
cols_to_clean <- c("On.board.Service", "Check.in.Service", "Seat.Comfort",
                   "In.flight.Service", "Gate.Location", "Cleanliness", 
                   "In.flight.Entertainment")

for (col in colnames(iv_data)) {
  if (is.factor(iv_data[[col]])) {
    iv_data[[col]] <- droplevels(iv_data[[col]])
  }
}

# Split the dataset into training and testing sets
proportion <- 0.7
split <- initial_split(iv_data, prop = proportion)
iv_training <- training(split)
iv_testing <- testing(split)

# Encode Satisfaction as binary (0 = Neutral or Dissatisfied, 1 = Satisfied)
iv_training$Satisfaction <- ifelse(iv_training$Satisfaction == "Satisfied", 1, 0)
iv_training$Satisfaction <- factor(iv_training$Satisfaction, levels = c(0, 1))
iv_testing$Satisfaction <- ifelse(iv_testing$Satisfaction == "Satisfied", 1, 0)
iv_testing$Satisfaction <- factor(iv_testing$Satisfaction, levels = c(0, 1))

# Using all data
model_log_iv <- glm(Satisfaction ~., 
               data = iv_training, family = "binomial")
summary(model_log_iv)

# Modelling
pred.test <- predict(model_log_iv, newdata = iv_testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
iv_testing$Satisfaction <- ifelse(iv_testing$Satisfaction == 1, "Satisfied", "Neutral or Dissatisfied")
iv_testing$Satisfaction <- factor(iv_testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

# Confusion matrix
confusionMatrix(pred.test.result, iv_testing$Satisfaction)

```

The new model with high information value variables shows a noticeable decrease in AIC value to 61591. This indicates that the model with high information value variables is better than the base model. The accuracy, sensitivity, and specificity values are also experiencing significant increase from the base model. This shows that there is a strong evidence that the new model with high information value variables are better than the base model.

Using the ANOVA test :

```{r}
anova(model_log, model_log_iv, test="Chisq")
```

Model IV with lower residual deviance and AIC value indicates that the model is better than the base model. The p-value is also very close to 0, which indicates that there is a significant difference between the two models. This means that the model with high information value variables is indeed better than the base model, confirming the result from summary function and the confusion matrix.

**Adjusting Threshold**

Although the High Information Value method already shows a better result than the base model, we want to check with other methods to see whether further improvement is possible. The third method that we will use to improve the model performance is by adjusting the threshold value. The default threshold value is 0.5, which means that if the predicted probability is greater than 0.5, the model will classify it as "Satisfied". However, we can adjust this threshold value to see if it improves the model performance. Based on the ROC curve, we can see that the optimal threshold value is actually around 0.468, not 0.5.

```{r}
coords(roc_obj, "best", ret = "threshold")
```

Now, we will run the confusion matrix again of the base model using the new threshold value of 0.468.

```{r}
all_testing <- testing
pred.test <- predict(model_log, newdata = testing, type = "response")

# Convert to class labels

pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
pred.test.result <- ifelse(pred.test > 0.4684071, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- ifelse(testing$Satisfaction == 1, "Satisfied", "Neutral or Dissatisfied")
all_testing$Satisfaction <- factor(all_testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, all_testing$Satisfaction)
```

The new threshold value indicates a slight increase in sensitivity value, while the specificity value decrease for about 1% than the previous model. Meanwhile, the accuracy score is relatively unchanged. This mixed result shows that the new threshold value method is not significantly better than the previous model.

**Dimensionality Reduction**

The last method that we will use to try improving the predictor model is by using PCA (Principal Component Analysis). PCA is a statistical technique used to reduce the dimensionality of a dataset while preserving as much variance as possible. It transforms the original variables into a new set of uncorrelated variables called principal components. The first few principal components capture most of the variance in the data, allowing us to reduce the number of variables while retaining important information.

```{r}
training_pca <- training %>%
  select(-Satisfaction) %>%
  mutate(across(everything(), ~ if (all(!is.na(as.numeric(as.character(.x))))) 
    as.numeric(as.character(.x)) else .x)) %>%
  select(where(is.numeric)) %>%
  scale(center = TRUE, scale = TRUE)

pca_results <- prcomp(training_pca)
summary(pca_results)
```

In this case, the PCA reached more than 90% of the variance with only 11 components. Now we can reduce the dimensionality of dataset and see if it improves the model performance.

```{r}
pca_data <- as.data.frame(pca_results$x[, 1:11])
pca_data$Satisfaction <- training$Satisfaction

model_log_pca <- glm(Satisfaction ~., 
               data = pca_data, family = "binomial")

summary(model_log_pca)
```

The summary of the PCA model shows that the AIC value increases to 81827, which is significantly higher than the base model AIC value of 70412. This indicates that the new model with PCA aprroach is likely to be worse than the base model. We will further confirm this finding by running the confusion matrix.

```{r}

train_center <- attr(training_pca, "scaled:center")
train_scale  <- attr(training_pca, "scaled:scale")

# 1. Convert test data to numeric and scale
test_pca_scaled <- testing %>%
  mutate(across(everything(), ~ if (all(!is.na(as.numeric(as.character(.x))))) 
    as.numeric(as.character(.x)) else NA_real_)) %>%  # forces fallback to numeric NA
  select(all_of(names(train_center))) %>%
  mutate(across(everything(), as.numeric)) %>%  # ensures everything is numeric
  scale(center = train_center, scale = train_scale)

# 2. Apply PCA transformation
test_pca <- as.data.frame(predict(pca_results, newdata = test_pca_scaled)[, 1:11])

# 3. Add target variable
test_pca$Satisfaction <- testing$Satisfaction

# 4. Predict using the model trained on PCA components
pred.test <- predict(model_log_pca, newdata = test_pca, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
test_pca$Satisfaction <- ifelse(test_pca$Satisfaction == 1, "Satisfied", "Neutral or Dissatisfied")
test_pca$Satisfaction <- factor(test_pca$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, test_pca$Satisfaction)
```

As we can see that all three important metrics (accuracy, sensitivity, and specificity) are significantly lower than the base model. This indicates that the PCA model is not better than the base model as indicated also by the ANOVA test.

# Interpretation <a name="4"></a>

## Result Visualization <a name="4.1"></a>

As a recap, we will again show the customer segmentation based on class using the chunk below. Economy and Economy Plus passengers are again grouped into Economy Combined, just as the model.

```{r passenger_segmentation_recap}

# Combine Economy and Economy Plus into a single category for visualization
class_summary <- data %>%
  mutate(Class = ifelse(Class %in% c("Economy", "Economy Plus"), "Economy Combined", Class)) %>%
  group_by(Class) %>%
  summarize(Count = n()) %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 1))  # Calculate percentages

# Plot the histogram for passenger segmentation with percentages
ggplot(class_summary, aes(x = reorder(Class, -Count), y = Count, fill = Class)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 4) +  # Add percentage labels
  labs(
    title = "Passenger Segmentation by Class",
    x = "Class",
    y = "Count",
    fill = "Class"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Business" = "royalblue", "Economy Combined" = "yellow2"))

remove(class_summary)
```

As seen above, our dataset is comprised of an almost equal proportion of Economy and Business class passengers. Each class represents 51.2% and 47.9% of the data respectively, with a total of 129,880 data points. Below, we show the satisfaction profiles of each customer segment.

```{r passenger_satisfaction_recap}

# Summarize satisfaction levels for each class
class_satisfaction_summary <- data %>%
  mutate(Class = ifelse(Class %in% c("Economy", "Economy Plus"), "Economy Combined", Class)) %>%
  group_by(Class, Satisfaction) %>%
  summarize(Count = n(), .groups = "drop") %>%
  mutate(Percentage = round(Count / sum(Count) * 100, 1))

# Plot satisfaction profiles with labels showing number and percentage
ggplot(class_satisfaction_summary, aes(x = Class, y = Percentage, fill = Satisfaction)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(Count, " (", Percentage, "%)")), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3.5) +  # Add labels above bars
  labs(
    title = "Passenger Class Segmentation and Satisfaction Profiles",
    x = "Passenger Class",
    y = "Percentage",
    fill = "Satisfaction Level"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Neutral or Dissatisfied" = "tomato2", "Satisfied" = "springgreen2"))

remove(class_satisfaction_summary)
```

In total, we see that only about 43.4% of passengers in our data were satisfied by their flight. Moreover, it is clearly evident that passengers of Economy class are much more frequently dissatisfied and make up the majority of the dissatisfaction. However, we also see that more than one third of business class passengers were dissatisfied with their experience. Clearly there is major room for improvement, hence we will answer the following question:

**What are the aspects of service that are important towards customer satisfaction?**

```{r important_aspects_of_service}

# Extract coefficients from the logistic regression models
coefs_business <- summary(model_log_business)$coefficients
coefs_economy <- summary(model_log_economy)$coefficients

# Create data frames for plotting
coef_df_business <- data.frame(
  Variable = rownames(coefs_business),
  Estimate = coefs_business[, "Estimate"]
)
coef_df_economy <- data.frame(
  Variable = rownames(coefs_economy),
  Estimate = coefs_economy[, "Estimate"]
)

# Remove the intercept for clarity
coef_df_business <- coef_df_business[coef_df_business$Variable != "(Intercept)", ]
coef_df_economy <- coef_df_economy[coef_df_economy$Variable != "(Intercept)", ]

# Select the top 5 most liked and least liked predictors for each class
top_5_liked_business <- coef_df_business[order(-coef_df_business$Estimate), ][1:5, ]
top_5_disliked_business <- coef_df_business[order(coef_df_business$Estimate), ][1:5, ]
top_5_liked_economy <- coef_df_economy[order(-coef_df_economy$Estimate), ][1:5, ]
top_5_disliked_economy <- coef_df_economy[order(coef_df_economy$Estimate), ][1:5, ]

# Combine liked and disliked for plotting
top_5_business <- rbind(
  transform(top_5_liked_business, Category = "Positive Influence"),
  transform(top_5_disliked_business, Category = "Negative Influence")
)
top_5_economy <- rbind(
  transform(top_5_liked_economy, Category = "Positive Influence"),
  transform(top_5_disliked_economy, Category = "Negative Influence")
)

# Plot the top 5 most liked and least liked predictors for Business Class
ggplot(top_5_business, aes(x = reorder(Variable, Estimate), y = Estimate, fill = Category)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top Influential Aspects (Business Class)",
       x = "Aspect of Service",
       y = "Coefficient Estimate (log-odds)") +
  scale_fill_manual(values = c("Positive Influence" = "royalblue", "Negative Influence" = "tomato")) +
  theme_minimal()

# Plot the top 5 most liked and least liked predictors for Economy Combined
ggplot(top_5_economy, aes(x = reorder(Variable, Estimate), y = Estimate, fill = Category)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top Influential Aspects (Economy Combined)",
       x = "Aspect of Service",
       y = "Coefficient Estimate (log-odds)") +
  scale_fill_manual(values = c("Positive Influence" = "yellow2", "Negative Influence" = "tomato")) +
  theme_minimal()
```

## Recommendation <a name="4.2"></a>
We have developed models to predict customer satisfaction based on various service aspects. Understanding satisfaction contributors is important because customers satisfaction mediates the impact of service quality on corporate braning and perceived value on brand loyalti (Hussain, 2016). The analysis revealed that the most significant predictors of satisfaction differ between Business and Economy class passengers. As we can see, the airline industry segmentation between business and economy class is quite strong, and the two classes have different preferences. Therefore, improving the satisfaction rate in business and economy segment requires different approaches.<br> <br>
As we can see from the model, there two types of top influences on customers satisfaction in economy class: **Positive Influence** and **Negative Influence**.<br> 
**Top Positive influence** refers to variables with the highest positive coefficients. These factors increase the likelihood of the desired outcome when improved.<br>
**Top Negative influence** refers to variables with the most negative coefficients. These factors decrease the likelihood of the desired outcome when increased or not addressed.However, this does not necessarily mean that improving these factors will directly bring the satisfaction goes down. It shows correlation but not causation. It indicates there are aspects in these factor that should be analyzed more deeply, those aspects might outweigh the benefit obtaines from improved factors.Thus, these factors should not be ignored<br><br>

<span style="color: #1E90FF; font-size: 18px;">**ECONOMY SEGMENT**</span><br>
Economy passengers shows high dissatisfaction level compared to business passengers at around 41.9% of total passengers. As we can see from the model, the most significant predictors of satisfaction for economy passengers are: <br> <br>
<ol>
<li>**In flight wifi service** : This shows the largest coefficient in the model indicating that satisfaction on this factor significantly contributes to overall customers' satisfaction. Enhancing WiFi services should be a top priority. Invest in high-speed, reliable internet and offer free or affordable connectivity options to economy-class passengers.
</li>
<li>**In flight entertainment**: Entertainment during flight matters for economy passengers. Increase option for entertainment by offering diverse movie, TV show, podcast, and music library. Therefore, ensuring personal screen for passengers and refresh the entertainment content regularly becomes important to drive the satisfaction</li>
<li>**Check in service**: Streamline check-in processes to make them faster and more efficient. Introduce self-service kiosks or mobile check-in apps to reduce wait times and improve customer satisfaction. The more efficient and user-friendly the check in, the higher satisfaction will be obtained from economy passengers</li>
<li>**Gate Location**: Make sure gates are easy to locate with reasonable distances. Economy passengers value this factor important, improving convenience in gate locating will improve satisfaction level. Airline could provide clear signage and information regarding the flight especially when the airport is relatively spacious with many gates, finding gaste then could be challenging</li>
<li>**On board service**: Improving on boarding service will drive the satisfaction from economy passengers. Upgrade the quality of food, beverages, and cabin crew responsiveness. A friendly and attentive crew can leave a lasting impression.</li>
</ol>
For the top negative influence factors, improving these factors might not directly bring the satisfaction up, but these factors should be analyzed more deeply. The mitigation for top negative influence factors are:<br> <br>
<ol>
<li>**Departure and Arrival time convenience**: Try to Survey passengers to align flight schedules with preferred times. Ensure more flexible options for travelers with varying needs.</li>
<li>**In flight service**:Identify specific areas of dissatisfaction through customer surveys and refine services accordingly. This could include meal options, cabin crew interactions, and overall service quality.</li>
<li>**Arrival Delay**: Try to Offer real-time updates to passengers when delays occur and compensate them with small gestures like snacks or discounts. Understanding more what aspects trigerring arrival delay that might dissatisfy passengers</li>
<li>**Ease of online booking**: Beside ensuring seamless experiences, try to find out what features in online booking that might not be favoured by passengers which could outweight the satisfaction level</li>
<li>**Arrival delay status on time**: If "on-time" flights have a negative impact due to unmet passenger expectations, consider adding services or gestures to acknowledge punctuality, like bonus miles or complimentary refreshments.</li>
</ol>
<br>
<span style="color: #1E90FF; font-size: 16px;">**BUSINESS SEGMENT**</span><br>
Business passengers shows higher satisfaction level compared to economy passenger. More passengers in business class are satisfied compared to those who are not satisfied. However, we need to take a look on satisfaction contributors to increse the satisfaction level. As we can see from the model, the most significant predictors of satisfaction for business passengers are: <br> <br>
<ol>
<li>**Online Boarding**: For business segment, simplify the booking process by improving website navigation and offering transparent pricing. Ensure mobile compatibility for seamless access. Allow more features on online boarding for example additional service that customers expect such as pick up, business loounge information. </li>
<li>**Leg room service**: Ensure ample legroom and ergonomic seating. Highlight this feature in marketing campaigns to attract business travelers. For business segment, this is one of their intention to choose the service, spacious room for their seating.</li>
<li>**On board service**: Train cabin crew to provide exceptional service tailored to business-class needs, such as personalized greetings and prompt assistance. </li>
<li>**Check in service**: Streamline check-in processes with dedicated counters or priority lanes for business-class passengers. Offer concierge services for added convenience. This is also the reason people choosing business service, putting attention to ensure seamless experience in check in service becomes matters. </li>
<li>**in flight entertainment**: Provide high-quality entertainment options, including noise-canceling headphones and curated content for business travelers. Airline could include games, various option of entertainment on screen, and make sure it is up to date. </li>
</ol>
<br>
In terms of top negative influence factors, improving these factors might not directly bring the satisfaction up, but these factors should be analyzed more deeply.Negative influence in model indicates that as the variable increases, it correlates with a decrease in satisfaction (or the desired outcome). However, it’s important to understand what this actually reflects.The mitigation for top negative influence factors are:<br> <br>
<ol>
<li>**Ease of online booking**:Identify usability issues and improve transparency in pricing and options. </li>
<li>**Departure delay status on-time**:Reduce delays through operational optimizations and offer compensatory perks when delays are unavoidable. </li>
<li>**Food and Drink**:Introduce premium and customizable meal options, catering to diverse dietary preferences. </li>
<li>**Arrival Delay**: Focus on punctual arrivals by optimizing flight schedules and improving ground operations. Provide real-time updates and assistance for delayed passengers. </li>
<li>**Departure and Arrival Convenience**:Optimize flight schedules to meet business travelers' expectations for flexibility and convenience </li>
</ol>
Addressing potential aspects that contribute to negative correlation from those factors could help to transform this to positive correlation. <br> <br>

<span style="color: #1E90FF; font-size: 16px;">**SUMMARY**</span><br>
Airline prioritize top contributor to satisfaction level of passengers.Improvement on this areas will bring positive impact on satisfaction level. In addition to this, airline should address challenges on top negative influence on satisfaction level. Identifying potential aspects that bring the negative impact on those factors. It help to increase the satisfaction level<br>
-For **Economy Segment**, Put in flight wifi service as top priority in business improvement, while progressing improvement on other four top factors such as in flight entertainment, check in service, gate location, and on board service. Focus to address arrival delay status and ease of online booking service, to identify contributor in negative influence<br>
- For **Business Segment**, Put online boarding as top priority in business improvement, while managing improvement on other top factors such as leg room service, on board service, check in service, and in flight entertainment.Focus to address departure and arrival time convenience as well as arrival delay, to identify contributor in negative influence <br><br>

## Further Improvement for Better Recommendation <a name="4.3"></a>

Although the current prediction model and approach has shown a robust result that can be used by airlines to improve their service, the predictor model can still be improved further. The current model is based on the separation between Economy and Business class passengers, which means that the model is only considering one dimension when predicting the most important variables that affect the passengers' satisfaction level. However, based on the survey data, there are still other dimensions that can be used to predict the customer satisfaction based on the survey variables. 

Among possible improvement area can be done by creating predictor model based on the combination of several dimensions that are easier to identify by the airlines, such as flight distance, which can be classified as short-haul, medium-haul, and long-haul flights, and age groups (children, youth, adult, and senior). Unlike variables like type of travel or customer type, flight distance and age groups can be quickly identified or classified by airlines without further questioning each passengers. Also, these two variables can be considered important for business decision making, such as in determining the flight route and the target market for each flight.

1.  **Flight Distance**:<br>
    -   Passengers flying in long-haul flights may have different expectations and experiences compared to those on short-haul flights, even though both type of flights can be served in both economy and business class. By segmenting the data based on flight distance, we can identify specific service aspects that are crucial for each group. Using this approach, airlines with long haul flights, such as Amsterdam to Jakarta, can capture a more detailed insights regarding the passengers' satisfaction point preference, which can be different from a shorter duration flight such as Amsterdam to London.<br>
2.  **Age Group**:<br>
    -   Different age groups may have varying preferences and expectations when it comes to airline services. By segmenting the data based on age, we can identify specific service aspects that are crucial for each group. For example, younger passengers may prioritize in-flight entertainment and WiFi connectivity, while older passengers may value comfort and service quality more. By understanding these differences, airlines can tailor their services to meet the needs of each age group. This is especially important for routes with popular tourist destination like the flight to Barcelona where the portion of younger passengers is likely to be higher (due to family holiday) than the flight without top tourist destinations. The airlines also can adjust the service priority during the holiday seasons to cover the increase in holiday-oriented passengers.<br>
    
The result of different set of most important variables based on flying class, flight distance, and age group can be used to create a more comprehensive recommendation for the airlines. This will help the airlines to improve their service quality and increase customer satisfaction by providing more data to support the decision-making process of airlines.

## Conclusion & Limitations <a name="4.4"></a>

**Conclusion**

Part 1 Project Background outlines our objective to predict whether future airline customers would be satisfied based on detailed aspects of the service they received, and to identify which service components most significantly influence satisfaction. To address this, we utilized the Airline Passenger Satisfaction dataset, consisting of 129,880 individual responses reflecting customer experiences across various service touchpoints.

Part 2 Data Exploration and Preparation involved gaining a deeper understanding of the data structure, relationships between variables, and the necessary steps for cleaning and transforming the data. This process ensured we could proceed with high-quality input for modeling. Important steps included handling missing values, treating outliers, creating meaningful feature groups (such as flight service categories), and transforming skewed variables. These efforts resulted in a dataset optimized for reliable analysis and insight extraction.

Part 3 Data Modeling focused on applying logistic regression and decision tree (C5.0) models to predict satisfaction and identify key influencing variables. Dimensionality reduction using PCA helped uncover underlying patterns, revealing components such as in-flight amenities, digital and booking experience, and flight punctuality. Both models demonstrated high predictive accuracy, with decision tree models capturing complex interactions and logistic regression offering interpretability of individual variable effects.

Part 4 Interpretation summarized the insights and provided actionable recommendations for stakeholders in the airline industry. The findings confirmed that in-flight service, seat comfort, cleanliness, food and drink, and on-board service are consistently the most impactful drivers of customer satisfaction. The hypothesis proposed in Part 1 is supported, highlighting that service quality—particularly during the flight itself—is central to satisfaction outcomes.

In conclusion, this analysis presents valuable insights for airline service providers seeking to enhance customer experience. By prioritizing core service components and operational reliability, airlines can not only improve satisfaction levels but also build stronger customer loyalty in a highly competitive market.

## References <a name=""></a>

Agresti, A. (2018). *Statistical thinking: Improving business performance* (2nd ed.). Boca Raton, FL: CRC Press.<br>

Biau, G. (2010). Analysis of a Random Forests Model. J. Mach. Learn. Res., 13, 1063-1095. <https://doi.org/10.5555/2503308.2343682><br> 

Federal Aviation Administration. (n.d.). Types of Delay. Retrieved April 15, 2025, from <https://aspm.faa.gov/aspmhelp/index/Types_of_Delay.html><br>

Hosmer, D. W., Lemeshow, S., & Sturdivant, R. X. (2013). *Applied logistic regression* (3rd ed.). Wiley<br>

Hussain, R. (2016). The mediating role of customer satisfaction: evidence from the airline industry. Asia Pacific Journal of Marketing and Logistics, 28, 234-255. https://doi.org/10.1108/APJML-01-2015-0001. <br>

International Air Transport Association. (2024, December). *Global outlook for air transport: December 2024*. IATA. Retrieved April 3, 2025 from <https://www.iata.org/en/iata-repository/publications/economic-reports/global-outlook-for-air-transport-december-2024/><br>

Landwehr, N., Hall, M., & Frank, E. (2003). Logistic Model Trees. Machine Learning, 59, 161-205. https://doi.org/10.1007/s10994-005-0466-3 <br>

Statista. (n.d.). *Flights - Shared Mobility*. Retrieved April 3, 2025, from <https://www.statista.com/outlook/mmo/shared-mobility/flights/worldwide> <br>

Wragg, David W., 1973. A Dictionary of Aviation. Reading, UK: Osprey Publishing <br>

Wright, M., & Ziegler, A. (2015). ranger: A Fast Implementation of Random Forests for High Dimensional Data in C++ and R. Journal of Statistical Software, 077, 1-17. <https://doi.org/10.18637/jss.v077.i01> <br>

Yan, K., & Hong, J. (2022). Adaptive cost-sensitive learning: Improving the convergence of intelligent diagnosis models under imbalanced data. Knowl. Based Syst., 241, 108296. <https://doi.org/10.1016/j.knosys.2022.108296>. <br>

Schwenk, H., & Bengio, Y. (2000). Boosting Neural Networks. Neural Computation, 12, 1869-1887. <https://doi.org/10.1162/089976600300015178>.<br> 